---
title: "Multivariate and Multi-pattern Tests with DTW"
bibliography: ../inst/REFERENCES.bib
output:
  pdf_document:
    number_sections: true
    toc: true
    toc_depth: 6
    df_print: kable
  md_document:
    toc: true
    toc_depth: 6
    df_print: kable
  html_document:
    number_sections: true
    toc: true
    toc_depth: 6
    toc_float: true
    df_print: kable
  word_document: default
---

In this notebook we want to conduct two different tests before we put it all together. First of all, we want to perform multivariate matching to see whether this has advantages over multiple univariate matches.

Then, we want to find out how `dtw` matches a query if it contains the reference more than once -- it is interesting to us which match is picked by dtw.

```{r warning=FALSE, message=FALSE}
library(dtw)
```

Load our common functions:

```{r}
source("./common-funcs.R", echo = FALSE)
```


```{r}
getAggStats <- function(dtwAlign, name, ref, query) {
  ex <- extract_signal_from_window(
    dtwAlign = dtwAlign, window = query,
    idxMethod = "smooth", smoothCnt = 3)
  
  f_ref <- pattern_approxfun(ref)
  f_query <- pattern_approxfun(query)
  
  return(cbind(
    data.frame(name = name),
    round(data.frame(
      area = area_diff_2_functions(f_ref, f_query)$value,
      KL_symm = stat_diff_2_functions_symmetric_KL(f_ref, f_query)$value,
      JSD = stat_diff_2_functions_symmetric_JSD(f_ref, f_query)$value,
      JSD_score = 1 - (stat_diff_2_functions_symmetric_JSD(f_ref, f_query)$value / log(2)),
      mono_rel = ex$monotonicity_rel,
      corr = stat_diff_2_functions_cor(f_ref, f_query)$value,
      sd = stat_diff_2_functions_sd(f_ref, f_query)$value,
      mae = stat_diff_2_functions_mae(f_ref, f_query)$value
    ), 5)
  ))
}
```


# Multiple scales

We want to measure how well `dtw` performs if the the reference/query are distorted (clinched or distorted). We will try linear distortions. Horizontal and vertical distortions are hence anti-proportional to each other, so that we will only manipulate along one axis in the following examples.

```{r}
set.seed(2)

ref1 <- density(rnorm(1000, mean = 3), n = 256)

ref1_signal <- list(
  x = 1:length(ref1$x),
  y = ref1$y / max(ref1$y)
)

ref1_f <- pattern_approxfun(ref1_signal$y)

plot(ref1_signal$y, ylim = c(0, 1))
```

Below is the "original query", which means that it is an example of how we would observe the signal in reality -- slightly different and probably with noise, also it is less stronger.

```{r}
set.seed(1)
q1_len <- 400
q1 <- sapply(seq(0, pi, len=q1_len), function(x) {
  .5 * (sin(x) * (1 - 1/(x + 1)) + runif(1, 0, 1e-1))
})

q1_signal <- list(
  x = 1:(q1_len + 2 * 106),
  y = c(runif(106,0,0.05), q1, runif(106, 0, 0.05))
)

q1_f <- pattern_approxfun(q1_signal$y)

plot(q1_signal, ylim = c(0,1))
```
In the following, we do some mostly linear scaling along the vertical axis.

```{r}
getAlign <- function(x){
  dtw::dtw(x = x, y = ref1_signal$y
           , step.pattern = dtw::asymmetric
           , keep = TRUE
           , open.begin = TRUE
           , open.end = TRUE)
}

q_normalized <- q1_signal$y / max(q1_signal$y)
q_times3 <-q1_signal$y * 3
q_times10 <-q1_signal$y * 10
q_half <- q1_signal$y / 2
q_exp <- 2.5^(3 + q1_signal$y) - 2.5^3
q_squared <- (q1_signal$y + 1)^3.5 - 1

align <- getAlign(x = q1_signal$y)
align_normalized <- getAlign(x = q_normalized)
align_times3 <- getAlign(x = q_times3)
align_times10 <- getAlign(x = q_times10)
align_half <- getAlign(x = q_half)
align_exp <- getAlign(x = q_exp)
align_squared <- getAlign(x = q_squared)

rbind(
  getAggStats(align, "original", ref1_signal$y, q1_signal$y),
  getAggStats(align_normalized, "normalized", ref1_signal$y, q_normalized),
  getAggStats(align_times3, "times 3", ref1_signal$y, q_times3),
  getAggStats(align_times10, "times 10", ref1_signal$y, q_times10),
  getAggStats(align_half, "half", ref1_signal$y, q_half),
  getAggStats(align_exp, "exp", ref1_signal$y, q_exp),
  getAggStats(align_squared, "squared", ref1_signal$y, q_squared)
)

plot(align, main = "original", type = "three")
plot(align_normalized, main = "normalized", type = "three")
plot(align_times3, main = "times 3", type = "three")
plot(align_times10, main = "times 10", type = "three")
plot(align_half, main = "half", type = "three")
plot(align_exp, main = "exp", type = "three")
plot(align_squared, main = "squared", type = "three")
```

From the above plots we learn one important lesson. Apparently, `dtw` does not handle cases well, where reference and query do not have the same co-domain. More specifically, the query's range should always be within (greater or equal the reference's minimum, less than or equal its maximum) the reference's range -- otherwise, `dtw` introduces plateaus/jumps, where parts of a signal cannot be matched.

We can see this in the first plot, the query is matched as long as it goes up. Then, however, there is a jump (not a plateau though) to where it starts decreasing, as the continued increase (or better: the values that are larger in the reference) cannot be matched in the query. Given how we perform matches of the curves, this case however is still good, as the its not a vertical jump, resulting in an extraction of the signal. The match would probably be good, too. However, the query is "smaller" than the reference, but still has a good resemblance -- monotonicity/continuity are good, as well as all stats based on the two function graphs (remember these are cut and scaled into the unit square).


# Multi-pattern Tests

```{r}
set.seed(3)

q2 <- c(
  rnorm(25, sd = .03),
  
  sapply(seq(0, pi, len=150), function(x) {
    sin(x) + runif(1, 0, .075)
  }),
  
  rnorm(50, sd = .02),
  
  sapply(seq(0, pi, length=70), function(x) {
    sin(x) * .8 + runif(1, 0, .04)
  }),
  
  rnorm(30, sd = .015)
)

# Make it fit into [0,1]
q2 <- q2 - min(q2)
q2 <- q2 / max(q2)

plot(q2)

```

Let's make some plots:

```{r}
q2_scale <- c(q2[1:199]*0.8, q2[200:325]*1.2)

align_q2 <- getAlign(x = q2)
align_q2_scale <- getAlign(x = q2_scale)
align_q2_scale_rev <- getAlign(x = rev(q2_scale))

plot(align_q2, type = "threeway")
plot(align_q2_scale, type = "threeway")
plot(align_q2_scale_rev, type = "threeway")
```

Some observations:

* `dtw` matches the curve/shape that will result in a lower cost (i.e., it does not necessarily match the curve that is spatially closer).
* Reversing a signal results in warping curve rotated by `180°`.
* We probably need to compute more metrics on the warping function than those we currently have implemented (see below).

## Additional metrics for the warping function

To further assess the goodness of match, we will match the warping function against its optimum: its own linear regression in the same interval. We have already done this to some degree to extract the warping-score (both absolute and relative).

However, we have developed powerful statistical tools to quantify the differences between two functions. So the goal is get the two functions, the warping function and its linear regression, and then extract the available metrics on those, too (anything from area, over KLD/JSD to correlation, rmse etc.).

```{r}
# The function `pattern_approxfun_warp` has been moved to `common-funcs.R`.
```

And here is an example:

```{r}
align_q2_warp <- pattern_approxfun_warp(align_q2)

plot_2_functions(align_q2_warp$f_warp, align_q2_warp$f_lm) + ggplot2::ggtitle("Warping function and its linear regression", "Both functions were scaled together and then put into the unit-square.")
plot_2_functions(align_q2_warp$f_warp_org, align_q2_warp$f_warp_np) + ggplot2::ggtitle("Warping function and its non-plateau version")
plot_2_functions(align_q2_warp$f_warp_org, align_q2_warp$f_warp_np_opt) + ggplot2::ggtitle("Warping function and its optimized non-plateau version")

area_diff_2_functions(align_q2_warp$f_warp, align_q2_warp$f_lm)$value
stat_diff_2_functions_cor(align_q2_warp$f_warp, align_q2_warp$f_lm)$value
stat_diff_2_functions_symmetric_JSD(align_q2_warp$f_warp, align_q2_warp$f_lm)$value
```

For any set of the above pair of functions, we can compute our metrics/statistics (e.g., area, correlation, KLD etc.). We developed a few alternatives because there is no one right answer to what is the theoretical optimum warping function for any arbitrary warping function.

In an ideal case, the whole query can be mapped to the reference, and it is mapped at a constant pace (i.e., no dilation or clinching). This would result in a perfect linear function with a slope in the range of $(0,\infty)$. However, we are using `dtw` and are expecting clinching and dilation -- in such a case, the optimal warping function is one that is strictly monotonically increasing (i.e., no plateaus), meaning that its intermittent slope may roam freely within $(0,\infty)$. In other words, a (full or partially non-)linear function that has a slope $> 0\, \forall\, X$, can still be optimal, considering that we are expecting dilation and clinching of the query.

Our solution to that is to extract those parts of the warping function that do in fact have a slope greater `0`. Then, we compare the original warping function to this one, as it one could argue that it represents the optimal warping function given the reference and the query (or, at least more optimal than the linear regression, as it resembles the warping function in greater detail). Also, by comparing we attempt to quantify the differences between the actual warping function, and a hypothetical optimum. One option is to compare against its linear regression. Another option is the one we just described. Also remember that we are already computing other metrics of the warping function, such as monotonicity/continuity, and the residuals of its linear regression (this is currently done in `extract_signal_from_window` but we may refactor this at a later point).




# Multivariate Tests



