---
title: "Technical Report: Final evaluation of Detecting the Fire Drill anti-pattern"
author: "Sebastian Hönel"
date: "`r format(Sys.time(), '%B %d, %Y')`"
bibliography: ../inst/REFERENCES.bib
urlcolor: blue
output:
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 6
    df_print: kable
    keep_tex: yes
  md_document:
    toc: yes
    toc_depth: 6
    df_print: kable
    variant: gfm
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 6
    toc_float: yes
    df_print: kable
  word_document: default
header-includes:
- \usepackage{bm}
- \usepackage{mathtools}
- \usepackage{xurl}
---

\newcommand*\mean[1]{\overline{#1}}
\newcommand{\norm}[1]{\left\lvert\,#1\,\right\rvert}
\newcommand{\infdiv}[2]{#1\;\|\;#2}
\newcommand\argmax[1]{\underset{#1}{arg\,max}}
\newcommand\argmin[1]{\underset{#1}{arg\,min}}


```{r setoptions, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
opts_chunk$set(tidy = TRUE, tidy.opts = list(indent=2))
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
source(file = "../helpers.R")
source(file = "./common-funcs.R")
source(file = "../models/modelsR6.R")
source(file = "../models/SRBTW-R6.R")

library(ggplot2)
library(ggpubr)
```

# Introduction\label{tr:fire-drill-final-eval-technical-report}

The purpose of this notebook to finally reconcile both of our approaches of detecting the Fire Drill anti-pattern, both using issue-tracking and source code data. The goal also to produce some consistent high-quality figures that will be used in the article.


# Tools

Here, we define some of the tools needed for this report.


## $m$-dimensional relative continuous Pearson sample correlation coefficient

First, here is the formula:

$$
\begin{aligned}
  f,g\;\dots&\,m\text{-dimensional continuous variables},
  \\[1ex]
  \bm{\mathit{S}}_f,\bm{\mathit{S}}_g=&\;a_{i,j},b_{i,j}\in \mathbb{R}^{m\times2}\,\text{, supports for}\;f,g\;\text{along each dimension,}
  \\[1ex]
  \overline{f}=\mathrm{E}[f]=&\;\left[\int_{a_{1,1}}^{a_{1,2}}\dots\int_{a_{m,1}}^{a_{m,2}}\,f(x_1,\dots,x_m)\,dx_1\dots dx_m\right]
  \\[0ex]
  &\;\;\times\prod_{i=1}^{m}\,(a_{i,1}-a_{i,2})^{-1}\,\text{, (equivalently for}\;\overline{g}\;\text{using}\;\bm{\mathit{S}}_g\text{),}
  \\[1em]
  \operatorname{corr}(f,g)=&\frac{
    \splitfrac{
        \Big(f\big(\mathsf{T}(a_{1,1},a_{1,2},x_1),\,\dots,\,\mathsf{T}(a_{m,1},a_{m,2},x_m)\big)-\mathrm{E}[f]\Big)
    }{
        \times\Big(g\big(\mathsf{T}(b_{1,1},b_{1,2},x_1),\,\dots,\,\mathsf{T}(b_{m,1},b_{m,2},x_m)\big)-\mathrm{E}[g]\Big)
    }
  }{
    \splitfrac{
      \sqrt{\left[\int_0^1\dots\int_0^1\,(f(\mathsf{T}(a_{1,1},a_{1,2},x_1),\,\dots)-\mathrm{E}[f])^2\,dx_1\,\dots\,dx_m\right]}
  }{
      \times\sqrt{\left[\int_0^1\dots\int_0^1\,(g(\mathsf{T}(b_{1,1},b_{1,2},x_1),\,\dots)-\mathrm{E}[g])^2\,dx_1\,\dots\,dx_m\right]}
    }
  }\,\text{,}\label{eq:mdim-corr}
  \\[1ex]
  c_{fg}=&\;\int_0^1\dots\int_0^1\,corr(f,g)\,dx_1\,\dots\,dx_m\,\text{.}
\end{aligned}
$$


However first, we will implement a continuous relative 1D version to test our approach. Let's generate some sample data that should be highly correlated, shown in figure \ref{fig:dens-example-data}. See how we deliberately have different means in order to dislocate both variables.


```{r}
set.seed(1)

a <- rnorm(500, mean = 5)
b <- rnorm(500, mean = 10)

dens_a <- density(a)
dens_b <- density(b)
dens_b$y <- -1 * dens_b$y

f_a <- approxfun(x = dens_a$x, y = dens_a$y, yleft = 0, yright = 0)
f_b <- approxfun(x = dens_b$x, y = dens_b$y, yleft = 0, yright = 0)
```

```{r dens-example-data, echo=FALSE, fig.cap="The densities of both normally-distributed variables.", fig.align="center", fig.pos="ht!"}
curve(f_a, min(c(dens_a$x, dens_b$x)), max(c(dens_a$x, dens_b$x)), ylim = range(c(dens_a$y, dens_b$y)))
curve(f_b, min(c(dens_a$x, dens_b$x)), max(c(dens_a$x, dens_b$x)), add = TRUE)
```

As expected, the spatial difference in location does not influence the correlation:

```{r}
cor(dens_a$y, dens_b$y)
```

Before we go further, we want to manually implement the coefficient and visualize the correlation vector. It is shown in figure \ref{fig:oned-corrvec}.

```{r}
corrvec_ab <- (dens_a$y - mean(dens_a$y)) * (dens_b$y - mean(dens_b$y)) /
              (sqrt(sum((dens_a$y - mean(dens_a$y))^2)) * sqrt(sum((dens_b$y - mean(dens_b$y))^2)))

sum(corrvec_ab)
```


```{r oned-corrvec, echo=FALSE, fig.cap="The one-dimensional correlation vector of the two discrete variables.", fig.align="center", fig.pos="ht!"}
plot(corrvec_ab)
```


### 1D continuous relative correlation

Now we define a continuous relative version of the Pearson sample correlation coefficient:

```{r}
coef_rel_pearson_1d <- function(f, g, supp_f = c(0,1), supp_g = c(0,1)) {
  # sum[ (x_i - bar_x) x (y_i - bar_y) ]
  # ------------------------------------
  # sqrt(sum[ (x_i - bar_x)^2 ]) x sqrt(...)
  
  transform_op <- function(a, b, x) a + x*b - x*a
  
  # Those work, too:
  # bar_f <- cubature::cubintegrate(
  #   f = f, lower = supp_f[1], upper = supp_f[2])$integral / (supp_f[2] - supp_f[1])
  # bar_g <- cubature::cubintegrate(
  #   f = g, lower = supp_g[1], upper = supp_g[2])$integral / (supp_g[2] - supp_g[1])
  
  bar_f <- cubature::cubintegrate(
    f = function(x) f(transform_op(supp_f[1], supp_f[2], x)), lower = 0, upper = 1)$integral
  bar_g <- cubature::cubintegrate(
    f = function(x) g(transform_op(supp_g[1], supp_g[2], x)), lower = 0, upper = 1)$integral
  
  denom_f <- sqrt(cubature::cubintegrate(f = function(x) {
    (f(transform_op(supp_f[1], supp_f[2], x)) - bar_f)^2
  }, lower = 0, upper = 1)$integral)
  denom_g <- sqrt(cubature::cubintegrate(f = function(x) {
    (g(transform_op(supp_g[1], supp_g[2], x)) - bar_g)^2
  }, lower = 0, upper = 1)$integral)
  
  fnum <- function(x) {
    (f(transform_op(a = supp_f[1], b = supp_f[2], x = x)) - bar_f) *
    (g(transform_op(a = supp_g[1], b = supp_g[2], x = x)) - bar_g)
  }
  
  numerator <- cubature::cubintegrate(f = fnum, lower = 0, upper = 1)$integral
  
  list(
    fnum = fnum,
    bar_f = bar_f,
    bar_g = bar_g,
    denom_f = denom_f,
    denom_g = denom_g,
    numerator = numerator,
    corr_func = function(x) {
      fnum(x) / (denom_f * denom_g)
    },
    corr_fg = numerator / (denom_f * denom_g)
  )
}
```

In figure \ref{fig:oned-corrfunc} we show the correlation of both continuous functions. Let's test using the previously generated data, density, and functions thereof:

```{r oned-corrfunc, echo=FALSE, fig.cap="The one-dimensional correlation function of the two continuous variables.", fig.align="center", fig.pos="ht!"}
temp <- coef_rel_pearson_1d(f = f_a, g = f_b, supp_f = range(dens_a$x), supp_g = range(dens_b$x))
temp

tempfff <- temp$corr_func
curve(tempfff, 0, 1)
```

We can see that the results are very similar, except for some decimals. Next, we compare some individual intermittent results, that should be very similar (discrete vs. continuous). Starting with the mean/expectation of either variable:

```{r}
c(mean(dens_a$y), mean(dens_b$y))
c(temp$bar_f, temp$bar_g)
```

Check, very similar. Next, we compute and plot the result of the numerator (cf. fig. \ref{fig:disc-num}):

```{r}
sum((dens_a$y - mean(dens_a$y)) * (dens_b$y - mean(dens_b$y)))
```

```{r disc-num, echo=FALSE, fig.cap="The discrete values that make up the numerator, plotted over all indexes.", fig.align="center", fig.pos="ht!"}
plot((dens_a$y - mean(dens_a$y)) * (dens_b$y - mean(dens_b$y)))
```

The continuous version of the numerator is in the previously computed result. If we integrate it, we get the __mean__ of the function (since the area under the curve over the interval $[0,1]$ is always a mean), not the sum. The function for the continuous numerator is shown in figure \ref{fig:cont-num}.

```{r cont-num, echo=FALSE, fig.cap="The relative function of the continuos numerator plotted over its support.", fig.align="center", fig.pos="ht!"}
tempfff <- temp$fnum
curve(tempfff, 0, 1)
```

In order to get the same result as we got from the previous summation, we need to multiply by the number of elements in the discrete variable. The following result is very close to what we got from the summation:

```{r}
cubature::cubintegrate(tempfff, 0, 1)$integral * length(a)
```

Let's check some values as computed in the denominator for the two variables (cf. fig. \ref{fig:disc-denoms}):

```{r}
c(
  sqrt(sum((dens_a$y - mean(dens_a$y))^2)),
  sqrt(sum((dens_b$y - mean(dens_b$y))^2)))
```


```{r disc-denoms, echo=FALSE, fig.cap="Plots of the continuous denominators for both data series.", fig.align="center", fig.pos="ht!"}
par(mfrow=c(2,1))
plot((dens_a$y - mean(dens_a$y))^2)
plot((dens_b$y - mean(dens_b$y))^2)
```

The continuous version of the denominators is the following (cf. fig. \ref{fig:cont-denoms}):

```{r}
tempf_a <- Vectorize(function(x) {
  transform_op <- function(a, b, x) a + x*b - x*a
  
  (f_a(transform_op(min(dens_a$x), max(dens_a$x), x = x)) - temp$bar_f)^2
})
tempf_b <- Vectorize(function(x) {
  transform_op <- function(a, b, x) a + x*b - x*a
  (f_b(transform_op(min(dens_b$x), max(dens_b$x), x = x)) - temp$bar_g)^2
})

c(
  sqrt(cubature::cubintegrate(tempf_a, 0, 1)$integral * length(a)),
  sqrt(cubature::cubintegrate(tempf_b, 0, 1)$integral * length(b)))
```


```{r cont-denoms, echo=FALSE, fig.cap="Plots of the continuous denominators for both data series.", fig.align="center", fig.pos="ht!"}
par(mfrow=c(2,1))
curve(tempf_a, 0, 1)
curve(tempf_b, 0, 1)
```

### 2D continuous relative correlation

Finally, we'll implement the 2D version. Note that we even allow the support of $y$ (the 2nd dimension) to depend on $x$. Therefore, we pass in the support for the 2nd dimension as function. The following function works well, but the number of evaluations should be limited to get results in time (e.g., $50$). I tried 1000 evaluations and got very precise results, but it ran for 30 minutes. With 50 evaluations, results are similarly close. Remember that we calculate correlations, and there it is often sufficient to have precision up to 2-3 decimals.

```{r}
coef_rel_pearson_2d <- function(f, g, supp_f_d1 = c(0,1), supp_g_d1 = c(0,1), supp_f_d2 = function(x) c(0,1), supp_g_d2 = function(x) c(0,1), maxEval = 50) {
  # sum[ (x_i - bar_x) x (y_i - bar_y) ]
  # ------------------------------------
  # sqrt(sum[ (x_i - bar_x)^2 ]) x sqrt(...)
  
  transform_op <- function(a, b, x) a + x*b - x*a
  
  double_int_mean <- function(func, supp_d1, supp_d2, maxEval = 50) {
    cubature::cubintegrate(f = function(x) {
      x1 <- transform_op(a = supp_d1[1], b = supp_d1[2], x = x)
      d2_a <- supp_d2(x1)[1]
      d2_b <- supp_d2(x1)[2]
      
      cubature::cubintegrate(f = function(y) {
        y1 <- transform_op(a = d2_a, b = d2_b, x = y)
        func(x1, y1)
      }, lower = 0, upper = 1, maxEval = maxEval)$integral
    }, lower = 0, upper = 1, maxEval = maxEval)$integral
  }
  
  bar_f <- double_int_mean(func = f, supp_d1 = supp_f_d1, supp_d2 = supp_f_d2, maxEval = maxEval)
  bar_g <- double_int_mean(func = g, supp_d1 = supp_g_d1, supp_d2 = supp_g_d2, maxEval = maxEval)
  
  
  denom_f <- sqrt(double_int_mean(func = function(x, y) {
    (f(x, y) - bar_f)^2
  }, supp_d1 = supp_f_d1, supp_d2 = supp_f_d2))
  denom_g <- sqrt(double_int_mean(func = function(x, y) {
    (g(x, y) - bar_g)^2
  }, supp_d1 = supp_g_d1, supp_d2 = supp_g_d2))
  
  fnum <- function(x, y) {
    x1_f <- transform_op(a = supp_f_d1[1], b = supp_f_d1[2], x = x)
    d2_f_a <- supp_f_d2(x1_f)[1]
    d2_f_b <- supp_f_d2(x1_f)[2]
    y1_f <- transform_op(a = d2_f_a, b = d2_f_b, x = y)
    
    x1_g <- transform_op(a = supp_g_d1[1], b = supp_g_d1[2], x = x)
    d2_g_a <- supp_g_d2(x1_g)[1]
    d2_g_b <- supp_g_d2(x1_g)[2]
    y1_g <- transform_op(a = d2_g_a, b = d2_g_b, x = y)
    
    (f(x1_f, y1_f) - bar_f) * (g(x1_g, y1_g) - bar_g)
  }
  
  numerator <- cubature::cubintegrate(f = function(x) {
    cubature::cubintegrate(f = function(y) {
      fnum(x, y)
    }, lower = 0, upper = 1, maxEval = maxEval)$integral
  }, lower = 0, upper = 1, maxEval = maxEval)$integral
  
  list(
    fnum = fnum,
    bar_f = bar_f,
    bar_g = bar_g,
    denom_f = denom_f,
    denom_g = denom_g,
    numerator = numerator,
    corr_func = function(x, y) {
      fnum(x, y) / (denom_f * denom_g)
    },
    corr_fg = numerator / (denom_f * denom_g)
  )
}
```

The correlation of these two two-dimensional variables is $\approx0.258$:

```{r}
CI_funcs <- readRDS(file = "../data/CI_p3avg_funcs.rds")
CI_req_p3avg <- CI_funcs$CI_req_p3avg
CI_dev_p3avg <- CI_funcs$CI_dev_p3avg

tempcorr <- loadResultsOrCompute(file = "../results/fd_fe_2dcorr.rds", computeExpr = {
  coef_rel_pearson_2db(f = CI_req_p3avg, g = CI_dev_p3avg, maxEval = 250)
})
tempcorr
```

In order to show the correlation in three dimensions, we'll compute a grid, cf. fig. \ref{fig:twod-correlation-req-dev-persp}:

```{r}
tempgrid <- loadResultsOrCompute(file = "../results/fd_fe_2dcorr_grid.rds", computeExpr = {
  outer(X = seq(0, 1, length.out = 75), Y = seq(0, 1, length.out = 75), FUN = tempcorr$fnum)
})
```

```{r twod-correlation-req-dev-persp, echo=FALSE, fig.cap="Correlation between the variables req\\% and dev\\%, plotted perspectively. On the left, we look at the correlation from above, while from the right, it is seen from underneath.", fig.align="center", fig.pos="ht!"}
suppressWarnings({
  par(mfrow=c(1,2))
  persp(tempgrid, zlim = c(-.03,.03), phi =  15, theta = -10, border = "#444444")
  persp(tempgrid, zlim = c(-.03,.03), phi = -15, theta = -10, border = "#444444")
})
```

From this example we see there is clearly both, positive and negative correlation. Here is the same as a colorful contour plot (fig. \ref{fig:twod-correlation-req-dev}) (correlation goes from blue/negative to red/postive, and no correlation is white):

```{r twod-correlation-req-dev, echo=FALSE, fig.cap="Correlation between the variables req\\% and dev\\%, plotted spatially.", fig.align="center", fig.pos="ht!"}
image(tempgrid, zlim = c(-.2,.2), col = colorRampPalette(colors = c("#0000ff", "#ffffff", "#ff0000"))(101)) # uneven number, so we have a true white for 0 correlation..
contour(tempgrid, zlim = c(-.2,.2), add = TRUE, col = "#444444")
```






















