---
title: "Technical Report: Final evaluation of Detecting the Fire Drill anti-pattern"
author: "Sebastian Hönel"
date: "`r format(Sys.time(), '%B %d, %Y')`"
bibliography: ../inst/REFERENCES.bib
urlcolor: blue
output:
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 6
    df_print: kable
    keep_tex: yes
  md_document:
    toc: yes
    toc_depth: 6
    df_print: kable
    variant: gfm
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 6
    toc_float: yes
    df_print: kable
  word_document: default
header-includes:
- \usepackage{bm}
- \usepackage{mathtools}
- \usepackage{xurl}
---

\newcommand*\mean[1]{\overline{#1}}
\newcommand{\norm}[1]{\left\lvert\,#1\,\right\rvert}
\newcommand{\infdiv}[2]{#1\;\|\;#2}
\newcommand\argmax[1]{\underset{#1}{arg\,max}}
\newcommand\argmin[1]{\underset{#1}{arg\,min}}


```{r setoptions, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
opts_chunk$set(tidy = TRUE, tidy.opts = list(indent=2))
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
source(file = "../helpers.R")
source(file = "./common-funcs.R")
source(file = "../models/modelsR6.R")
source(file = "../models/SRBTW-R6.R")

library(ggplot2)
library(ggpubr)
library(reshape2)
library(Cairo)
```

```{r}
library(extrafont)
font_import()
loadfonts()
```


# Introduction\label{tr:fire-drill-final-eval-technical-report}

The purpose of this notebook to finally reconcile both of our approaches of detecting the Fire Drill anti-pattern, both using issue-tracking and source code data. The goal also to produce some consistent high-quality figures that will be used in the article.


# Tools

Here, we define some of the tools needed for this report.


## Theme for figures

We will try to only use `ggplot2`-figures, but we want them to look as plain as the built-in plot-functionality. The template we are currently using has a textwidth of $4.68596$ inches, so we'll floor to two decimals. We define a function to create and return some custom styles. Note that in the rendered report, some figures are not shown as the sections are not evaluated or echoed, because they are meant to be run manually. Therefore, you'll have to check out the source code of this notebook in some cases.

```{r echo=FALSE}
par(family = "LM Roman 10")

fe_fig_width <- 4.68
fe_alpha <- 0.9

fe_gg2_style <- function(facets = FALSE) {
  sty <- theme_minimal(base_size = 10) + theme(
    text = element_text(family = "serif"),
    axis.title.x.bottom = element_text(margin = margin(t = 8)),
    axis.title.y.left = element_text(margin = margin(r = 8)))
  
  if (facets) {
    sty <- sty + theme(
      text = element_text(family = "LM Roman 10"),
      panel.spacing = unit(1, "mm"),
      panel.border = element_rect(color = "#d8d8d8", fill = NA, size = .1),
      strip.background = element_rect(fill="#f2f2f2", color = "#d8d8d8"),
      strip.text.x = element_text(color="black", margin = margin(t = 2, b = 2)),
      strip.text.y = element_text(color="black", margin = margin(l = 4, r = 2)))
  }
  sty
}
```


```{r echo=FALSE}
saveTikzAndEps <- function(ggInst, name, width = fe_fig_width, height = 3.2) {
  tikzDevice::tikz(file = paste0("../figures/fig_fdfe_", name, ".tex"), width = width, height = height, verbose = FALSE)
  print(ggInst)
  dev.off()
  
  saveAndPlotAsEPS(ggplotInstance = ggInst, fileName = paste0("../figures/fig_fdfe_", name, ".eps"), width = width, height = height)
}

saveTikzExpr <- function(expr, name, width = fe_fig_width, height = 3.2) {
  tikzDevice::tikz(file = paste0("../figures/fig_fdfe_", name, ".tex"), width = width, height = height, verbose = FALSE)
  eval(expr)
  dev.off()
}

saveEPSExpr <- function(expr, name, width = fe_fig_width, height = 3.2) {
  setEPS()
  postscript(paste0("../figures/fig_fdfe_", name, ".eps"), width = width, height = height)
  eval(expr)
  dev.off()
}

#' Add text with background box to a plot
#'
#' \code{boxtext} places a text given in the vector \code{labels} 
#' onto a plot in the base graphics system and places a coloured box behind 
#' it to make it stand out from the background.
#' 
#' @param x numeric vector of x-coordinates where the text labels should be 
#' written. If the length of \code{x} and \code{y} differs, the shorter one 
#' is recycled.
#' @param y numeric vector of y-coordinates where the text labels should be 
#' written. 
#' @param labels a character vector specifying the text to be written.
#' @param col.text the colour of the text 
#' @param col.bg color(s) to fill or shade the rectangle(s) with. The default 
#' \code{NA} means do not fill, i.e., draw transparent rectangles.
#' @param border.bg color(s) for rectangle border(s). The default \code{NA}
#' omits borders. 
#' @param adj one or two values in [0, 1] which specify the x (and optionally 
#' y) adjustment of the labels. 
#' @param pos a position specifier for the text. If specified this overrides 
#' any adj value given. Values of 1, 2, 3 and 4, respectively indicate 
#' positions below, to the left of, above and to the right of the specified 
#' coordinates.
#' @param offset when \code{pos} is specified, this value gives the offset of 
#' the label from the specified coordinate in fractions of a character width.
#' @param padding factor used for the padding of the box around 
#' the text. Padding is specified in fractions of a character width. If a 
#' vector of length two is specified then different factors are used for the
#' padding in x- and y-direction.    
#' @param cex numeric character expansion factor; multiplied by 
#' code{par("cex")} yields the final character size. 
#' @param font the font to be used
#'
#' @return Returns the coordinates of the background rectangle(s). If 
#' multiple labels are placed in a vactor then the coordinates are returned
#' as a matrix with columns corresponding to xleft, xright, ybottom, ytop. 
#' If just one label is placed, the coordinates are returned as a vector.
#' @author Ian Kopacka
#' @examples
#' ## Create noisy background
#' plot(x = runif(1000), y = runif(1000), type = "p", pch = 16, 
#' col = "#40404060")
#' boxtext(x = 0.5, y = 0.5, labels = "some Text", col.bg = "#b2f4f480", 
#'     pos = 4, font = 2, cex = 1.3, padding = 1)
#' @export
boxtext <- function(x, y, labels = NA, col.text = NULL, col.bg = NA, 
        border.bg = NA, adj = NULL, pos = NULL, offset = 0.5, 
        padding = c(0.5, 0.5), cex = 1, font = graphics::par('font')){

    ## The Character expansion factro to be used:
    theCex <- graphics::par('cex')*cex

    ## Is y provided:
    if (missing(y)) y <- x

    ## Recycle coords if necessary:    
    if (length(x) != length(y)){
        lx <- length(x)
        ly <- length(y)
        if (lx > ly){
            y <- rep(y, ceiling(lx/ly))[1:lx]           
        } else {
            x <- rep(x, ceiling(ly/lx))[1:ly]
        }       
    }

    ## Width and height of text
    textHeight <- graphics::strheight(labels, cex = theCex, font = font)
    textWidth <- graphics::strwidth(labels, cex = theCex, font = font)

    ## Width of one character:
    charWidth <- graphics::strwidth("e", cex = theCex, font = font)

    ## Is 'adj' of length 1 or 2?
    if (!is.null(adj)){
        if (length(adj == 1)){
            adj <- c(adj[1], 0.5)            
        }        
    } else {
        adj <- c(0.5, 0.5)
    }

    ## Is 'pos' specified?
    if (!is.null(pos)){
        if (pos == 1){
            adj <- c(0.5, 1)
            offsetVec <- c(0, -offset*charWidth)
        } else if (pos == 2){
            adj <- c(1, 0.5)
            offsetVec <- c(-offset*charWidth, 0)
        } else if (pos == 3){
            adj <- c(0.5, 0)
            offsetVec <- c(0, offset*charWidth)
        } else if (pos == 4){
            adj <- c(0, 0.5)
            offsetVec <- c(offset*charWidth, 0)
        } else {
            stop('Invalid argument pos')
        }       
    } else {
      offsetVec <- c(0, 0)
    }

    ## Padding for boxes:
    if (length(padding) == 1){
        padding <- c(padding[1], padding[1])
    }

    ## Midpoints for text:
    xMid <- x + (-adj[1] + 1/2)*textWidth + offsetVec[1]
    yMid <- y + (-adj[2] + 1/2)*textHeight + offsetVec[2]

    ## Draw rectangles:
    rectWidth <- textWidth + 2*padding[1]*charWidth
    rectHeight <- textHeight + 2*padding[2]*charWidth    
    graphics::rect(xleft = xMid - rectWidth/2, 
            ybottom = yMid - rectHeight/2, 
            xright = xMid + rectWidth/2, 
            ytop = yMid + rectHeight/2,
            col = col.bg, border = border.bg)

    ## Place the text:
    graphics::text(xMid, yMid, labels, col = col.text, cex = theCex, font = font, 
            adj = c(0.5, 0.5))    

    ## Return value:
    if (length(xMid) == 1){
        invisible(c(xMid - rectWidth/2, xMid + rectWidth/2, yMid - rectHeight/2,
                        yMid + rectHeight/2))
    } else {
        invisible(cbind(xMid - rectWidth/2, xMid + rectWidth/2, yMid - rectHeight/2,
                        yMid + rectHeight/2))
    }    
}
```



## $m$-dimensional relative continuous Pearson sample correlation coefficient

First, here is the formula:

$$
\begin{aligned}
  f,g\;\dots&\,m\text{-dimensional continuous variables},
  \\[1ex]
  \bm{\mathit{S}}_f,\bm{\mathit{S}}_g=&\;a_{i,j},b_{i,j}\in \mathbb{R}^{m\times2}\,\text{, supports for}\;f,g\;\text{along each dimension,}
  \\[1ex]
  \overline{f}=\mathrm{E}[f]=&\;\left[\int_{a_{1,1}}^{a_{1,2}}\dots\int_{a_{m,1}}^{a_{m,2}}\,f(x_1,\dots,x_m)\,dx_1\dots dx_m\right]
  \\[0ex]
  &\;\;\times\prod_{i=1}^{m}\,(a_{i,1}-a_{i,2})^{-1}\,\text{, (equivalently for}\;\overline{g}\;\text{using}\;\bm{\mathit{S}}_g\text{),}
  \\[1ex]
  =&\;\int_0^1\dots\int_0^1\,f(\mathsf{T}(a_{1,1},a_{1,2},x_1),\,\dots\,,\mathsf{T}(a_{m,1},a_{m,2},x_m))\,dx_1\,\dots\,dx_m\,\text{,}
  \\[1em]
  \operatorname{corr}(f,g)=&\frac{
    \splitfrac{
        \Big(f\big(\mathsf{T}(a_{1,1},a_{1,2},x_1),\,\dots,\,\mathsf{T}(a_{m,1},a_{m,2},x_m)\big)-\mathrm{E}[f]\Big)
    }{
        \times\Big(g\big(\mathsf{T}(b_{1,1},b_{1,2},x_1),\,\dots,\,\mathsf{T}(b_{m,1},b_{m,2},x_m)\big)-\mathrm{E}[g]\Big)
    }
  }{
    \splitfrac{
      \left[\int_0^1\dots\int_0^1\,\Big(f\big(\mathsf{T}(a_{1,1},a_{1,2},x_1),\,\dots\big)-\mathrm{E}[f]\Big)^2\,dx_1\,\dots\,dx_m\right]^{\frac{1}{2}}
  }{
      \times\left[\int_0^1\dots\int_0^1\,\Big(g\big(\mathsf{T}(b_{1,1},b_{1,2},x_1),\,\dots\big)-\mathrm{E}[g]\Big)^2\,dx_1\,\dots\,dx_m\right]^{\frac{1}{2}}
    }
  }\,\text{,}\label{eq:mdim-corr}
  \\[1em]
  c_{fg}=&\;\int_0^1\dots\int_0^1\,corr(f,g)\,dx_1\,\dots\,dx_m\,\text{.}
\end{aligned}
$$


However first, we will implement a continuous relative 1D version to test our approach. Let's generate some sample data that should be highly correlated, shown in figure \ref{fig:dens-example-data}. See how we deliberately have different means in order to dislocate both variables.


```{r}
set.seed(1)

a <- rnorm(500, mean = 5)
b <- rnorm(500, mean = 10)

dens_a <- density(a)
dens_b <- density(b)
dens_b$y <- -1 * dens_b$y

f_a <- approxfun(x = dens_a$x, y = dens_a$y, yleft = 0, yright = 0)
f_b <- approxfun(x = dens_b$x, y = dens_b$y, yleft = 0, yright = 0)
```

```{r dens-example-data, echo=FALSE, fig.height=4, fig.cap="The densities of both normally-distributed variables.", fig.align="center", fig.pos="ht!"}
curve(f_a, min(c(dens_a$x, dens_b$x)), max(c(dens_a$x, dens_b$x)), ylim = range(c(dens_a$y, dens_b$y)))
curve(f_b, min(c(dens_a$x, dens_b$x)), max(c(dens_a$x, dens_b$x)), add = TRUE)
```

As expected, the spatial difference in location does not influence the correlation:

```{r}
cor(dens_a$y, dens_b$y)
```

Before we go further, we want to manually implement the coefficient and visualize the correlation vector. It is shown in figure \ref{fig:oned-corrvec}.

```{r}
corrvec_ab <- (dens_a$y - mean(dens_a$y)) * (dens_b$y - mean(dens_b$y)) /
              (sqrt(sum((dens_a$y - mean(dens_a$y))^2)) * sqrt(sum((dens_b$y - mean(dens_b$y))^2)))

sum(corrvec_ab)
```


```{r oned-corrvec, echo=FALSE, fig.height=3, fig.cap="The one-dimensional correlation vector of the two discrete variables.", fig.align="center", fig.pos="ht!"}
plot(corrvec_ab)
```


### 1D continuous relative correlation

Now we define a continuous relative version of the Pearson sample correlation coefficient:

```{r}
coef_rel_pearson_1d <- function(f, g, supp_f = c(0,1), supp_g = c(0,1)) {
  # sum[ (x_i - bar_x) x (y_i - bar_y) ]
  # ------------------------------------
  # sqrt(sum[ (x_i - bar_x)^2 ]) x sqrt(...)
  
  transform_op <- function(a, b, x) a + x*b - x*a
  
  # Those work, too:
  # bar_f <- cubature::cubintegrate(
  #   f = f, lower = supp_f[1], upper = supp_f[2])$integral / (supp_f[2] - supp_f[1])
  # bar_g <- cubature::cubintegrate(
  #   f = g, lower = supp_g[1], upper = supp_g[2])$integral / (supp_g[2] - supp_g[1])
  
  bar_f <- cubature::cubintegrate(
    f = function(x) f(transform_op(supp_f[1], supp_f[2], x)), lower = 0, upper = 1)$integral
  bar_g <- cubature::cubintegrate(
    f = function(x) g(transform_op(supp_g[1], supp_g[2], x)), lower = 0, upper = 1)$integral
  
  denom_f <- sqrt(cubature::cubintegrate(f = function(x) {
    (f(transform_op(supp_f[1], supp_f[2], x)) - bar_f)^2
  }, lower = 0, upper = 1)$integral)
  denom_g <- sqrt(cubature::cubintegrate(f = function(x) {
    (g(transform_op(supp_g[1], supp_g[2], x)) - bar_g)^2
  }, lower = 0, upper = 1)$integral)
  
  fnum <- function(x) {
    (f(transform_op(a = supp_f[1], b = supp_f[2], x = x)) - bar_f) *
    (g(transform_op(a = supp_g[1], b = supp_g[2], x = x)) - bar_g)
  }
  
  numerator <- cubature::cubintegrate(f = fnum, lower = 0, upper = 1)$integral
  
  list(
    fnum = fnum,
    bar_f = bar_f,
    bar_g = bar_g,
    denom_f = denom_f,
    denom_g = denom_g,
    numerator = numerator,
    corr_func = function(x) {
      fnum(x) / (denom_f * denom_g)
    },
    corr_fg = numerator / (denom_f * denom_g)
  )
}
```

In figure \ref{fig:oned-corrfunc} we show the correlation of both continuous functions. Let's test using the previously generated data, density, and functions thereof:

```{r oned-corrfunc, echo=FALSE, fig.height=3, fig.cap="The one-dimensional correlation function of the two continuous variables.", fig.align="center", fig.pos="ht!"}
temp <- coef_rel_pearson_1d(f = f_a, g = f_b, supp_f = range(dens_a$x), supp_g = range(dens_b$x))
temp

tempfff <- temp$corr_func
curve(tempfff, 0, 1)
```

We can see that the results are very similar, except for some decimals. Next, we compare some individual intermittent results, that should be very similar (discrete vs. continuous). Starting with the mean/expectation of either variable:

```{r}
c(mean(dens_a$y), mean(dens_b$y))
c(temp$bar_f, temp$bar_g)
```

Check, very similar. Next, we compute and plot the result of the numerator (cf. fig. \ref{fig:disc-num}):

```{r}
sum((dens_a$y - mean(dens_a$y)) * (dens_b$y - mean(dens_b$y)))
```

```{r disc-num, echo=FALSE, fig.height=3, fig.cap="The discrete values that make up the numerator, plotted over all indexes.", fig.align="center", fig.pos="ht!"}
plot((dens_a$y - mean(dens_a$y)) * (dens_b$y - mean(dens_b$y)))
```

The continuous version of the numerator is in the previously computed result. If we integrate it, we get the __mean__ of the function (since the area under the curve over the interval $[0,1]$ is always a mean), not the sum. The function for the continuous numerator is shown in figure \ref{fig:cont-num}.

```{r cont-num, echo=FALSE, fig.height=3, fig.cap="The relative function of the continuos numerator plotted over its support.", fig.align="center", fig.pos="ht!"}
tempfff <- temp$fnum
curve(tempfff, 0, 1)
```

In order to get the same result as we got from the previous summation, we need to multiply by the number of elements in the discrete variable. The following result is very close to what we got from the summation:

```{r}
cubature::cubintegrate(tempfff, 0, 1)$integral * length(a)
```

Let's check some values as computed in the denominator for the two variables (cf. fig. \ref{fig:disc-denoms}):

```{r}
c(
  sqrt(sum((dens_a$y - mean(dens_a$y))^2)),
  sqrt(sum((dens_b$y - mean(dens_b$y))^2)))
```


```{r disc-denoms, echo=FALSE, fig.height=3, fig.cap="Plots of the discrete denominators for both data series.", fig.align="center", fig.pos="ht!"}
par(mfrow=c(1,2))
plot((dens_a$y - mean(dens_a$y))^2)
plot((dens_b$y - mean(dens_b$y))^2)
```

The continuous version of the denominators is the following (cf. fig. \ref{fig:cont-denoms}):

```{r}
tempf_a <- Vectorize(function(x) {
  transform_op <- function(a, b, x) a + x*b - x*a
  
  (f_a(transform_op(min(dens_a$x), max(dens_a$x), x = x)) - temp$bar_f)^2
})
tempf_b <- Vectorize(function(x) {
  transform_op <- function(a, b, x) a + x*b - x*a
  (f_b(transform_op(min(dens_b$x), max(dens_b$x), x = x)) - temp$bar_g)^2
})

c(
  sqrt(cubature::cubintegrate(tempf_a, 0, 1)$integral * length(a)),
  sqrt(cubature::cubintegrate(tempf_b, 0, 1)$integral * length(b)))
```


```{r cont-denoms, echo=FALSE, fig.height=3, fig.cap="Plots of the continuous denominators for both data series.", fig.align="center", fig.pos="ht!"}
par(mfrow=c(1,2))
curve(tempf_a, 0, 1)
curve(tempf_b, 0, 1)
```

### 2D continuous relative correlation

Finally, we'll implement the 2D version. Note that we even allow the support of $y$ (the 2nd dimension) to depend on $x$. Therefore, we pass in the support for the 2nd dimension as function. The following function works well, but the number of evaluations should be limited to get results in time (e.g., $50$). I tried 1000 evaluations and got very precise results, but it ran for 30 minutes. With 50 evaluations, results are similarly close. Remember that we calculate correlations, and there it is often sufficient to have precision up to 2-3 decimals.

```{r}
coef_rel_pearson_2d <- function(f, g, supp_f_d1 = c(0,1), supp_g_d1 = c(0,1), supp_f_d2 = function(x) c(0,1), supp_g_d2 = function(x) c(0,1), maxEval = 50) {
  # sum[ (x_i - bar_x) x (y_i - bar_y) ]
  # ------------------------------------
  # sqrt(sum[ (x_i - bar_x)^2 ]) x sqrt(...)
  
  transform_op <- function(a, b, x) a + x*b - x*a
  
  double_int_mean <- function(func, supp_d1, supp_d2, maxEval = 50) {
    cubature::cubintegrate(f = function(x) {
      x1 <- transform_op(a = supp_d1[1], b = supp_d1[2], x = x)
      d2_a <- supp_d2(x1)[1]
      d2_b <- supp_d2(x1)[2]
      
      cubature::cubintegrate(f = function(y) {
        y1 <- transform_op(a = d2_a, b = d2_b, x = y)
        func(x1, y1)
      }, lower = 0, upper = 1, maxEval = maxEval)$integral
    }, lower = 0, upper = 1, maxEval = maxEval)$integral
  }
  
  bar_f <- double_int_mean(func = f, supp_d1 = supp_f_d1, supp_d2 = supp_f_d2, maxEval = maxEval)
  bar_g <- double_int_mean(func = g, supp_d1 = supp_g_d1, supp_d2 = supp_g_d2, maxEval = maxEval)
  
  
  denom_f <- sqrt(double_int_mean(func = function(x, y) {
    (f(x, y) - bar_f)^2
  }, supp_d1 = supp_f_d1, supp_d2 = supp_f_d2))
  denom_g <- sqrt(double_int_mean(func = function(x, y) {
    (g(x, y) - bar_g)^2
  }, supp_d1 = supp_g_d1, supp_d2 = supp_g_d2))
  
  fnum <- function(x, y) {
    x1_f <- transform_op(a = supp_f_d1[1], b = supp_f_d1[2], x = x)
    d2_f_a <- supp_f_d2(x1_f)[1]
    d2_f_b <- supp_f_d2(x1_f)[2]
    y1_f <- transform_op(a = d2_f_a, b = d2_f_b, x = y)
    
    x1_g <- transform_op(a = supp_g_d1[1], b = supp_g_d1[2], x = x)
    d2_g_a <- supp_g_d2(x1_g)[1]
    d2_g_b <- supp_g_d2(x1_g)[2]
    y1_g <- transform_op(a = d2_g_a, b = d2_g_b, x = y)
    
    (f(x1_f, y1_f) - bar_f) * (g(x1_g, y1_g) - bar_g)
  }
  
  numerator <- pracma::integral2(fun = fnum, xmin = 0, xmax = 1, ymin = 0, ymax = 1)$Q
  
#  numerator <- cubature::cubintegrate(f = function(x) {
#    cubature::cubintegrate(f = function(y) {
#      fnum(x, y)
#    }, lower = 0, upper = 1, maxEval = maxEval)$integral
#  }, lower = 0, upper = 1, maxEval = maxEval)$integral
  
  list(
    fnum = fnum,
    bar_f = bar_f,
    bar_g = bar_g,
    denom_f = denom_f,
    denom_g = denom_g,
    numerator = numerator,
    corr_func = function(x, y) {
      fnum(x, y) / (denom_f * denom_g)
    },
    corr_fg = numerator / (denom_f * denom_g)
  )
}
```

The correlation of these two two-dimensional variables is $\approx0.258$:

```{r}
CI_funcs <- readRDS(file = "../data/CI_p3avg_funcs.rds")
CI_req_p3avg <- CI_funcs$CI_req_p3avg
CI_dev_p3avg <- CI_funcs$CI_dev_p3avg

tempcorr <- loadResultsOrCompute(file = "../results/fdfe_2dcorr.rds", computeExpr = {
  coef_rel_pearson_2d(f = CI_req_p3avg, g = CI_dev_p3avg, maxEval = 250)
})
tempcorr
```

In order to show the correlation in three dimensions, we'll compute a grid, cf. fig. \ref{fig:twod-correlation-req-dev-persp}:

```{r}
tempgrid <- loadResultsOrCompute(file = "../results/fdfe_2dcorr_grid.rds", computeExpr = {
  outer(X = seq(0, 1, length.out = 75), Y = seq(0, 1, length.out = 75), FUN = tempcorr$fnum)
})
```

```{r twod-correlation-req-dev-persp, echo=FALSE, fig.cap="Correlation between the variables req\\% and dev\\%, plotted perspectively. On the left, we look at the correlation from above, while from the right, it is seen from underneath.", fig.align="center", fig.pos="ht!"}
suppressWarnings({
  par(mfrow=c(1,2))
  persp(tempgrid, zlim = c(-.1,.1), phi =  15, theta = -10, border = "#444444")
  persp(tempgrid, zlim = c(-.1,.1), phi = -15, theta = -10, border = "#444444")
})
```

From this example we see there is clearly both, positive and negative correlation. Here is the same as a colorful contour plot (fig. \ref{fig:twod-correlation-req-dev}) (correlation goes from blue/negative to red/positive, and no correlation is white):

```{r twod-correlation-req-dev, echo=FALSE, fig.cap="Correlation between the variables req\\% and dev\\%, plotted spatially.", fig.align="center", fig.pos="ht!"}

plotexpr <- expression({
  par(mar=c(4,4,.85,.85), family = "LM Roman 10")
  
  fields::image.plot(
    tempgrid,
    zlim = .75*c(-max(tempgrid), max(tempgrid)),
    legend.shrink = .5, legend.lab = "Relative correlation strength", legend.line = 2.5,
    col = colorRampPalette(colors = c("#0000ff", "#ffffff", "#ff0000"))(101)) # uneven number, so we have a true white for 0 correlation..
  title(xlab = "Relative project time", ylab = "Average confidence", line = 2.5)
  grid()
  contour(tempgrid, zlim = .75*c(-max(tempgrid), max(tempgrid)), add = TRUE, col = "#888888", drawlabels = FALSE)
})
eval(plotexpr)
```

```{r eval=FALSE, echo=FALSE}
saveTikzExpr(expr = plotexpr, name = "2d_corr_req_dev_ci", height = 2.5, width = fe_fig_width)
```


```{r eval=FALSE, echo=FALSE}
pdf(file = "../figures/fig_fdfe_2d_corr_req_dev_ci.pdf", width = fe_fig_width, height = 2.5, pointsize = 8)
eval(plotexpr)
dev.off()
```


```{r echo=FALSE, eval=FALSE}
temp <- melt(tempgrid)
temp$Var1 <- (temp$Var1 - min(temp$Var1)) / (max(temp$Var1) - min(temp$Var1))
temp$Var2 <- (temp$Var2 - min(temp$Var2)) / (max(temp$Var2) - min(temp$Var2))
ggtemp <- ggplot(data = temp, aes(x = Var1, y = Var2, z = value, fill = value)) +
  geom_tile(aes(fill = value), alpha = fe_alpha) +
  geom_contour(color = "grey") +
  scale_fill_gradientn(
    colours = colorRampPalette(colors = c("#3333dd", "#ffffff", "#dd3333"))(101),
    limits = .5*c(-max(tempgrid), max(tempgrid))) +
  labs(x = "Relative time", y = "Average confidence", fill = "Relative correlation strength") +
  fe_gg2_style() +
  theme(
    legend.key.height = unit(2, "mm"),
    legend.key.width = unit(1, "cm"),
    legend.position = "bottom",
    legend.title = element_text(margin = margin(t=-15, r=3)))
ggtemp
```

```{r echo=FALSE, eval=FALSE}
invisible(saveTikzAndEps(ggtemp, "2d_corr_req_dev_ci", height = 3.2))
```


# Plots

## Type-I patterns

For reproducing plots it is easiest to just sample from the functions.

```{r echo=FALSE, eval=FALSE}
tempdf <- data.frame(matrix(ncol = 4, nrow = 0))
templ <- list(
  REQ = req, DEV = dev, DESC = desc,
  REQ_CI_LOWER = req_ci_lower, REQ_CI_UPPER = req_ci_upper
)
n <- 500
x <- seq(from = 0, to = 1, length.out = n)

for (name in names(templ)) {
  tempdf <- rbind(tempdf, data.frame(
    x = x,
    y = templ[[name]](x),
    v = name,
    p = if (grepl("R", name)) "P1" else "P2"
  ))
}
```

```{r echo=FALSE, eval=FALSE}
tempdf_rib <- rbind(data.frame(
  x = x,
  ymin = req_ci_lower(x),
  ymax = req_ci_upper(x),
  p = "P1"
), data.frame(
  x = x,
  ymin = rep(0, length(x)),
  ymax = dev_ci_upper(x),
  p = "P2"
))

```


```{r echo=FALSE, eval=FALSE}
ggplot(data = tempdf, mapping = aes(x = x, y = y, color = v)) +
  geom_line() +
  geom_ribbon(data = tempdf_rib, mapping = aes(x = x, y = 1, ymin = ymin, ymax = ymax, fill = p), alpha = 0.4, size = 0, show.legend = FALSE, inherit.aes = FALSE) +
  fe_gg2_style(facets = FALSE) +
  scale_fill_manual(values = c("red", "blue")) +
  scale_linetype_manual(values = c(rep("solid", 2), rep("twodashed", 3))) +
  facet_wrap(p ~.)
```


```{r echo=FALSE, eval=FALSE}
plotexpr <- rlang::expr({
  par(mfrow=c(2,1), mar=c(3,3,.25,0), family = "LM Roman 10")#, oma=rep(0,4))


  plot(x=0, y=0, xlim = c(0,1), ylim = c(0,1), col = "#00000000",
       xlab = "", ylab = "")
  title(ylab = "Relative cum. time spent", line = 2)
  grid()
  
  polygon(
    x = seq(from = 0, to = 1, length.out = 50),
    y = req_ci_upper(seq(from = 0, to = 1, length.out = 50)),
    col = "#ff000033",
    border = NA)
  curve(req_ci_lower, 0, 1, col = "red", lty = 2, add = TRUE)
  curve(req_ci_upper, 0, 1, col = "red", lty = 3, add = TRUE)
  curve(req, 0, 1, col = "red", lty = 1, lwd = 2, add = TRUE)
  
  for (i in 1:length(all_signals)) {
    pId <- names(all_signals)[i]
    tempf <- all_signals[[pId]]$REQ$get0Function()
    curve(tempf, 0, 1, col="#dd8888", lty=2, add=TRUE)
  }
  
  # other thresholds:
  abline(h = req(t_1), col = "#888888")
  text(x = .82, y = .64, paste0("$\\mathrm{req}(t_1)=", round(req(t_1), 2), "$"))
  
  abline(h = req_ci_upper(t_1), col = "#888888")
  #text(x = .28, y = .86, paste0("$CI_{req}^{+}(t_1)=", round(req_ci_upper(t_1), 2), "$"))
  boxtext(x = .28, y = .86, col.bg = "white", padding = c(0.1, 1.5), labels = paste0("$\\mathrm{CI}_{\\mathrm{req}}^{+}(t_1)=", round(req_ci_upper(t_1), 2), "$"))
  
  abline(v = 0, col = "#888888")
  abline(v = 1, col = "#888888")
  
  # t_1:
  abline(v = t_1)
  text(x = .47, y = .04, paste0("$t_1=", t_1, "$"))
  
  legend(.74, .42, legend = c("req", "upper CI", "lower CI"),
         col = "red", lty = c(1, 2, 3), lwd = c(2, 1, 1),
         bg = "white", cex = .85)
  
  
  
  
  plot(x=0, y=0, xlim = c(0,1), ylim = c(0,1), col = "#00000000",
       xlab = "", ylab = "")
  title(xlab = "Relative project time",
        ylab = "Relative cum. time spent", line = 2)
  grid()
  
  polygon(
    x = c(seq(from = 0, to = 1, length.out = 50), 1),
    y = c(dev_ci_upper(seq(from = 0, to = 1, length.out = 50)), 0),
    col = "#0000ff33",
    border = NA)
  
  curve(dev_ci_upper, 0, 1, col = "blue", lty = 3, add = TRUE)
  curve(desc, 0, 1, add = TRUE, col = "forestgreen", lty = 1, lwd = 3)
  curve(dev, 0, 1, col = "blue", lty = 1, lwd = 3, add = TRUE)
  
  for (i in 1:length(all_signals)) {
    pId <- names(all_signals)[i]
    tempf <- all_signals[[pId]]$DEV$get0Function()
    curve(tempf, 0, 1, col="#8888dd", lty=2, add=TRUE)
    tempf <- all_signals[[pId]]$DESC$get0Function()
    curve(tempf, 0, 1, col="#8888dd", lty=2, add=TRUE)
  }
  
  # t_1(dev), t_2(dev):
  abline(h = dev(t_1), col = "#888888")
  #text(x = .2, y = .2, paste0("$\\mathrm{dev}(t_1)=", round(dev(t_1), 3), "$"))
  boxtext(x = .12, y = .2, col.bg = "white", padding = c(0.1, 1.5), labels = paste0("$\\mathrm{dev}(t_1)=", round(dev(t_1), 3), "$"))
  abline(h = dev(t_2), col = "#888888")
  #text(x = .51, y = .46, paste0("$DEV(t_2)=", round(dev(t_2), 2), "$"))
  boxtext(x = .51, y = .46, col.bg = "white", padding = c(0.1, 1.5), labels = paste0("$\\mathrm{dev}(t_2)=", round(dev(t_2), 2), "$"))
  
  # other thresholds:
  abline(h = dev_ci_upper(t_2), col = "#888888")
  #text(x = .73, y = .77, paste0("$CI_{dev}^{+}(t_2)=", round(dev_ci_upper(t_2), 2), "$"))
  boxtext(x = .73, y = .77, col.bg = "white", padding = c(0.1, 1.5), labels = paste0("$\\mathrm{CI}_{\\mathrm{dev}}^{+}(t_2)=", round(dev_ci_upper(t_2), 2), "$"))
  abline(h = desc(1), col = "#888888")
  text(x = .87, y = .2, paste0("$\\mathrm{desc}(1)=", round(desc(1), 2), "$"))
  
  
  abline(v = 0, col = "#888888")
  abline(v = 1, col = "#888888")
  
  # t_1,t_2:
  abline(v = t_1)
  #text(x = .47, y = .98, paste0("$t_1=", t_1, "$"))
  abline(v = t_2)
  text(x = .775, y = .98, paste0("$t_2=", t_2, "$"))
  
  legend(.01, 1.01, legend = c("dev upper CI", "dev", "desc"),
         col = c("blue", "blue", "forestgreen"), lty = c(3, 1, 1),
         lwd = c(1, 2, 2), bg = "white", cex = .85)
})

eval(plotexpr)
```


```{r echo=FALSE, eval=FALSE}
saveTikzExpr(name = "model_t1_it", height = 4.2, width = 3.3, expr = plotexpr)
# saveEPSExpr(name = "model_t1_it", height = 4.2, expr = plotexpr)
```


```{r}
plotexpr <- rlang::expr({
  layout(mat = matrix(data = c(1,2), nrow = 2), heights = c(4,5), widths = c(1,1))
  par(mar=c(3,3,1,0), family = "LM Roman 10")
  
  
  plot(x=0, y=0, xlim = c(0,1), ylim = c(0,1.02), col = "#00000000",
       xlab = "", ylab = "", cex.axis = 0.8)
  title(ylab = "Amount of activity", line = 2)
  grid()
  
  curve2(p1_signals$A$get0Function(), 0, 1, col = "red", lty = 1, lwd = 2, add = TRUE)
  curve2(p1_signals$CP$get0Function(), 0, 1, col = "blue", lty = 1, lwd = 2, add = TRUE)
  curve2(p1_signals$FREQ$get0Function(), 0, 1, col = "forestgreen", lty = 1, lwd = 2, add = TRUE)
  curve2(p1_signals$SCD$get0Function(), 0, 1, col = "purple", lty = 1, lwd = 2, add = TRUE)
  
  abline(v = fd_data_boundaries[1], col = "#888888")
  text(x = fd_data_boundaries[1], y = 1.14, paste0("$b_1=", fd_data_boundaries[1], "$"), cex = .8, xpd = NA)
  abline(v = fd_data_boundaries[2], col = "#888888")
  text(x = fd_data_boundaries[2], y = 1.14, paste0("$b_2=", fd_data_boundaries[2], "$"), cex = .8, xpd = NA)
  abline(v = fd_data_boundaries[3], col = "#888888")
  text(x = fd_data_boundaries[3], y = 1.14, paste0("$b_3=", fd_data_boundaries[3], "$"), cex = .8, xpd = NA)
  
  legend(.18, 1.03, legend = c("$\\mathsf{A}$", "$\\mathsf{CP}$", "$\\mathsf{FREQ}$", "$\\mathsf{SCD}$"),
         col = c("red", "blue", "forestgreen", "purple"), lty = 1, lwd = 2,
         bg = "white", cex = .6)
  
  


  plot(x=0, y=0, xlim = c(0,1), ylim = c(0,1), col = "#00000000",
       xlab = "", ylab = "", cex.axis = 0.8)
  title(xlab = "Relative project time",
        ylab = "Relative cum. time spent", line = 2, cex = 0.8)
  grid()
  
  polygon(
    x = seq(from = 0, to = 1, length.out = 50),
    y = req_ci_upper(seq(from = 0, to = 1, length.out = 50)),
    col = "#ff000033",
    border = NA)
  
  polygon(
    x = c(seq(from = 0, to = 1, length.out = 50), 1),
    y = c(dev_ci_upper(seq(from = 0, to = 1, length.out = 50)), 0),
    col = "#0000ff33",
    border = NA)
  
  curve(req_ci_lower, 0, 1, col = "red", lty = 2, add = TRUE)
  curve(req_ci_upper, 0, 1, col = "red", lty = 3, add = TRUE)
  curve(req, 0, 1, col = "red", lty = 1, lwd = 2, add = TRUE)
  
  # other thresholds:
  #abline(h = req(t_1), col = "#888888")
  #text(x = .57, y = .64, paste0("$\\mathsf{REQ}(t_1)=", round(req(t_1), 2), "$"), cex = .8)
  
  #abline(h = req_ci_upper(t_1), col = "#888888")
  #text(x = .575, y = .86, labels = paste0("$\\mathsf{CI}_{\\mathsf{REQ}}^{+}(t_1)=", round(req_ci_upper(t_1), 2), "$"), cex = .8)
  
  #abline(v = 0, col = "#888888")
  #abline(v = 1, col = "#888888")
  
  # t_1:
  abline(v = t_1)
  text(x = t_1, y = 1.1, paste0("$t_1=", t_1, "$"), cex = .8, xpd = NA)
  
  curve(dev_ci_upper, 0, 1, col = "blue", lty = 3, add = TRUE)
  curve(desc, 0, 1, add = TRUE, col = "forestgreen", lty = 1, lwd = 3)
  curve(dev, 0, 1, col = "blue", lty = 1, lwd = 3, add = TRUE)
  
  # t_1(dev), t_2(dev):
  #abline(h = dev(t_1), col = "#888888")
  #text(x = .27, y = .2, labels = paste0("$\\mathsf{CI}_{\\mathsf{DEV}}^{+}(t_1))=", round(dev_ci_upper(t_1), 3), "$"), cex = .8)
  #abline(h = dev(t_2), col = "#888888")
  #text(x = .695, y = .45, labels = paste0("$\\mathsf{DEV}(t_2)=", round(dev(t_2), 2), "$"), cex = .8)
  
  # other thresholds:
  #abline(h = dev_ci_upper(t_2), col = "#888888")
  #text(x = .69, y = .75, labels = paste0("$\\mathsf{CI}_{\\mathsf{DEV}}^{+}(t_2)=", round(dev_ci_upper(t_2), 2), "$"), cex = .8)
  #abline(h = desc(1), col = "#888888")
  #text(x = .86, y = .2, paste0("$\\mathsf{DESC}(1)=", round(desc(1), 2), "$"), cex = .8)
  
  # t_2:
  abline(v = t_2)
  text(x = t_2, y = 1.1, paste0("$t_2=", t_2, "$"), cex = .8, xpd = NA)
  
  legend(.01, 1.02, legend = c("$\\mathsf{REQ}$", "$\\mathsf{CI}_{\\mathsf{REQ}}^{+}$", "$\\mathsf{CI}_{\\mathsf{REQ}}^{-}$", "$\\mathsf{DESC}$", "$\\mathsf{DEV}$", "$\\mathsf{CI}_{\\mathsf{DEV}}^{+}$"),
         col = c(rep("red", 3), "forestgreen", "blue", "blue"), lty = c(1, 3, 2, 1, 1, 3), lwd = c(2, 1, 1, 2, 2, 1),
         bg = "white", cex = .6)
})

eval(plotexpr)
```


```{r echo=FALSE, eval=FALSE}
saveTikzExpr(name = "model_t1_comb", height = 3.5, expr = plotexpr, width = 3)
```





```{r echo=FALSE, eval=FALSE}
plotexpr <- rlang::expr({
  par(mfrow=c(2,1), mar=c(3,3,.25,0), family = "LM Roman 10")#, family = "serif")#, oma=rep(0,4))


  plot(x=0, y=0, xlim = c(0,1), ylim = c(0,1), col = "#00000000",
       xlab = "", ylab = "")
  title(ylab = "Relative cum. time spent", line = 2)
  grid()
  
  polygon(
    x = c(seq(from = 0, to = 1, length.out = 50), t1_wavg),
    y = c(req_ci_upper_p2a(seq(from = 0, to = 1, length.out = 50)), req_ci_lower_p2a(t1_wavg)),
    col = "#ff00000d",
    border = NA)
  curve(req_p2a, 0, 1, col="red", lwd=2, add = TRUE)
  curve(req_ci_upper_p2a, 0, 1, col="red", lty=3, lwd=1.5, add = TRUE)
  curve(req_ci_lower_p2a, 0, 1, col="red", lty=2, lwd=1.5, add = TRUE)
  
  for (i in 1:length(all_signals)) {
    pId <- names(all_signals)[i]
    tempf <- all_signals[[pId]]$REQ$get0Function()
    curve(tempf, 0, 1, col="#dd8888", lty=2, add=TRUE)
  }
  
  abline(v = 0, col = "#888888")
  abline(v = 1, col = "#888888")
  
  # t_1:
  abline(v = t1_wavg)
  text(x = .47, y = .04, paste0("$\\hat{t}_1=", round(t1_wavg, 2), "$"))
  abline(v = t_1, col = "#aaaaaa")
  text(x = .32, y = .97, paste0("$t_1=", t_1, "$"))
  
  legend(.74, .42, legend = c("upper CI", "lower CI", "REQ"),
         col = "red", lty = c(3, 2, 1), lwd = c(1, 1, 2),
         bg = "white", cex = .85)
  
  
  
  
  plot(x=0, y=0, xlim = c(0,1), ylim = c(0,1), col = "#00000000",
       xlab = "", ylab = "")
  title(xlab = "Relative project time",
        ylab = "Relative cum. time spent", line = 2)
  grid()
  
  polygon(
    x = c(seq(from = 0, to = 1, length.out = 50), 1),
    y = c(dev_ci_upper_p2a(seq(from = 0, to = 1, length.out = 50)), 0),
    col = "#0000ff0d",
    border = NA)
  curve(dev_p2a, 0, 1, col="blue", lwd=2, add = TRUE)
  curve(desc_p2a, 0, 1, col="forestgreen", lwd=2, add = TRUE)
  tempf <- function(x) rep(0, length(x))
  curve(tempf, 0, 1, col="blue", lty=2, lwd=1.5, add = TRUE)
  curve(dev_ci_upper_p2a, 0, 1, col="blue", lty=3, lwd=1.5, add = TRUE)
  
  for (i in 1:length(all_signals)) {
    pId <- names(all_signals)[i]
    tempf <- all_signals[[pId]]$DEV$get0Function()
    curve(tempf, 0, 1, col="#8888dd", lty=2, add=TRUE)
    tempf <- all_signals[[pId]]$DESC$get0Function()
    curve(tempf, 0, 1, col="#55bb55", lty=2, add=TRUE)
  }
  
  abline(v = 0, col = "#888888")
  abline(v = 1, col = "#888888")
  
  abline(v = t2_wavg)
  text(x = .5, y = .98, paste0("$\\hat{t}_2=", round(t2_wavg, 2), "$"))
  abline(v = t_2, col = "#aaaaaa")
  text(x = .92, y = .32, paste0("$t_2=", t_2, "$"))
  
  legend(.01, 1.01, legend = c("DEV upper CI", "DEV", "DESC"),
         col = c("blue", "blue", "forestgreen"), cex = .85,
         lty = c(3, 1, 1), lwd = c(1, 2, 2), bg = "white")
})

eval(plotexpr)
```

```{r echo=FALSE, eval=FALSE}
saveTikzExpr(name = "model_t2_it", height = 4.2, expr = plotexpr)
saveEPSExpr(name = "model_t2_it", height = 4.2, expr = plotexpr)
```



```{r echo=FALSE, eval=FALSE}
plotexpr <- rlang::expr({
  par(mfrow=c(2,1), mar=c(3,3,.25,0), family = "LM Roman 10")#, oma=rep(0,4))


  plot(x=0, y=0, xlim = c(0,1), ylim = c(0,1), col = "#00000000",
       xlab = "", ylab = "")
  title(ylab = "Value", line = 2)
  grid()
  
  curve2(p1_signals$A$get0Function(), 0, 1, col = "red", lty = 1, lwd = 2, add = TRUE)
  curve2(p1_signals$CP$get0Function(), 0, 1, col = "blue", lty = 1, lwd = 2, add = TRUE)
  curve2(p1_signals$FREQ$get0Function(), 0, 1, col = "forestgreen", lty = 1, lwd = 2, add = TRUE)
  curve2(p1_signals$SCD$get0Function(), 0, 1, col = "purple", lty = 1, lwd = 2, add = TRUE)
  
  for (i in 1:length(project_signals)) {
    pId <- names(project_signals)[i]
    tempf <- project_signals[[pId]]$A$get0Function()
    curve2(tempf, 0, 1, col="#dd8888", lty=2, add=TRUE)
    # tempf <- project_signals[[pId]]$CP$get0Function()
    # curve2(tempf, 0, 1, col="#8888dd", lty=2, add=TRUE)
    # tempf <- project_signals[[pId]]$FREQ$get0Function()
    # curve2(tempf, 0, 1, col="#88bb88", lty=2, add=TRUE)
    # tempf <- project_signals[[pId]]$SCD$get0Function()
    # curve2(tempf, 0, 1, col="#dd88dd", lty=2, add=TRUE)
  }
  
  abline(v = fd_data_boundaries[1], col = "#888888")
  text(x = .165, y = .45, paste0("$b_1=", fd_data_boundaries[1], "$"))
  abline(v = fd_data_boundaries[2], col = "#888888")
  text(x = .54, y = .98, paste0("$b_2=", fd_data_boundaries[2], "$"))
  abline(v = fd_data_boundaries[3], col = "#888888")
  text(x = .795, y = .31, paste0("$b_3=", fd_data_boundaries[3], "$"))
  
  abline(v = 0, col = "#888888")
  abline(v = 1, col = "#888888")
  
  
  
  
  plot(x=0, y=0, xlim = c(0,1), ylim = c(0,1), col = "#00000000",
       xlab = "", ylab = "")
  title(ylab = "Value", xlab = "Relative project time", line = 2)
  grid()
  
  curve2(p2_signals$A$get0Function(), 0, 1, col = "red", lty = 1, lwd = 2, add = TRUE)
  curve2(p2_signals$CP$get0Function(), 0, 1, col = "blue", lty = 1, lwd = 2, add = TRUE)
  curve2(p2_signals$FREQ$get0Function(), 0, 1, col = "forestgreen", lty = 1, lwd = 2, add = TRUE)
  curve2(p2_signals$SCD$get0Function(), 0, 1, col = "purple", lty = 1, lwd = 2, add = TRUE)
  
  abline(v = p2_best_thetaB[2], col = "#888888")
  text(x = .16, y = .3, paste0("$\\hat{b}_1=", round(p2_best_thetaB[2], 3), "$"))
  abline(v = p2_best_thetaB[3], col = "#888888")
  text(x = .345, y = .97, paste0("$\\hat{b}_2=", round(p2_best_thetaB[3], 3), "$"))
  abline(v = p2_best_thetaB[4], col = "#888888")
  text(x = .59, y = .97, paste0("$\\hat{b}_3=", round(p2_best_thetaB[4], 3), "$"))
  
  legend(.28, .73, legend = c("A", "CP", "FREQ", "SCD"),
         col = c("red", "blue", "forestgreen", "purple"), lty = 1, lwd = 2,
         bg = "white", cex = .85)
})

eval(plotexpr)
```


```{r echo=FALSE, eval=FALSE}
saveTikzExpr(expr = plotexpr, name = "model_t1_t2_sc", height = 4.2)
```




```{r echo=FALSE, eval=FALSE}
tempdf <- data.frame(matrix(ncol = 4, nrow = 0))
colnames(tempdf) <- c("x", "y", "p", "v")
n <- 500
x <- seq(from = 0, to = 1, length.out = n)
for (pId in names(all_signals)) {
  for (v in c("REQ", "DEV", "DESC")) {
    tempdf <- rbind(tempdf, data.frame(
      x = x,
      y = sapply(X = x, FUN = all_signals[[pId]][[v]]$get0Function()),
      p = pId,
      v = v
    ))
  }
}
```

```{r echo=FALSE, eval=FALSE}
plot_all_req_dev <- ggplot(data = tempdf, aes(x = x, y = y, color = v)) +
  geom_line(size = 1/3) +
  facet_wrap(p ~., labeller = as_labeller(`names<-`(paste0("Project ", 1:9), paste0("Project", 1:9)))) +
  fe_gg2_style(facets = TRUE) +
  labs(color = "Variable") + xlab("Relative project time") + ylab("Relative cumulative time spent") +
  scale_color_manual(values = c("forestgreen", "blue", "red")) +
  scale_x_continuous(breaks = c(0, .25, .5, .75, 1), labels = paste0(c(0, .25, .5, .75, 1))) +
  theme(legend.position = "bottom", legend.margin = margin(t=-7.5))

saveTikzAndEps(ggInst = plot_all_req_dev, name = "all_projects_it", height = 3.4)
```


```{r echo=FALSE, eval=FALSE}
tempdf <- data.frame(matrix(ncol = 4, nrow = 0))
colnames(tempdf) <- c("x", "y", "p", "v")

n <- 500
x <- seq(from = 0, to = 1, length.out = n)
for (pId in levels(student_projects$project)) {
  for (v in c("A", "CP", "FREQ", "SCD")) {
    tempdf <- rbind(tempdf, data.frame(
      x = x,
      y = sapply(X = x, FUN = project_signals[[pId]][[v]]$get0Function()),
      p = pId,
      v = v
    ))
  }
}
```


```{r echo=FALSE, eval=FALSE}
plot_all_acp <- ggplot(data = tempdf, aes(x = x, y = y, color = v)) +
  geom_line(size = 1/3) +
  facet_wrap(p ~., labeller = as_labeller(`names<-`(paste0("Project ", 1:9), levels(student_projects$project)))) +
  fe_gg2_style(facets = TRUE) +
  labs(color = "Variable") + xlab("Relative project time") + ylab("Value") +
  theme(legend.position = "bottom", legend.margin = margin(t=-7.5)) +
  scale_x_continuous(breaks = c(0, .25, .5, .75, 1), labels = paste0(c(0, .25, .5, .75, 1))) +
  scale_color_manual(values = c("red", "blue", "forestgreen", "purple"))

saveTikzAndEps(ggInst = plot_all_acp, name = "all_projects_sc", height = 3.4)
```



```{r echo=FALSE, eval=FALSE}
z_req <- loadResultsOrCompute(file = "../results/ci_p3avg_z_req.rds", computeExpr = {
  compute_z_p3avg(varname = "REQ", x = x, y = y)
})
z_dev <- loadResultsOrCompute(file = "../results/ci_p3avg_z_dev.rds", computeExpr = {
  compute_z_p3avg(varname = "DEV", x = x, y = y)
})
z_desc <- loadResultsOrCompute(file = "../results/ci_p3avg_z_desc.rds", computeExpr = {
  compute_z_p3avg(varname = "DESC", x = x, y = y)
})

plotexpr <- rlang::expr({
  par(mfrow=c(3,1), mar=c(4,4,.8,.8), family = "LM Roman 10", cex.axis = 1.5)#, oma=rep(0,4))
  
  tempz <- fields::interp.surface.grid(obj = z_req, grid.list = list(
    x = seq(from = 0, to = 1, length.out = 60), y = seq(from = 0, to = 1, length.out = 80)))
  fields::image.plot(x = tempz$x, y = tempz$y, z = tempz$z,
    xlab = "", ylab = "",
    col = colorRampPalette(colors = c("#ffffff", "#ff6666"))(100), zlim = c(0,1),
    legend.shrink = .5, legend.lab = "Confidence", legend.line = 3)#, legend.cex = 1)
  title(ylab = "Relative cum. time spent", line = 3, cex.lab = 1.5)
  grid()
  
  curve(req_p3, 0, 1, col = "red", add = TRUE)
  
  curve(req_ci_upper_p3avg, 0, 1, col = "#ff000033", add = TRUE, lty = 2)
  curve(req_ci_lower_p3avg, 0, 1, col = "#ff000033", add = TRUE, lty = 3)
  
  for (i in 1:length(all_signals)) {
    pId <- names(all_signals)[i]
    tempf <- all_signals[[pId]]$REQ$get0Function()
    curve(tempf, 0, 1, col="#00000033", lty=if (omega[i] == 0) 3 else 2, add=TRUE)
  }
  
  legend(0.65, .37, legend = c("REQ (avg.)", "Weight $=0$", "Weight $>0$"), bg = "white",
         col = c("red", rep("#00000033", 2)), lty = c(1, 3, 2), lwd = c(2.5, 2, 2), cex = 1.3)
  
  
  
  tempz <- fields::interp.surface.grid(obj = z_dev, grid.list = list(
    x = seq(from = 0, to = 1, length.out = 60), y = seq(from = 0, to = 1, length.out = 80)))
  fields::image.plot(x = tempz$x, y = tempz$y, z = tempz$z,
    xlab = "", ylab = "",
    col = colorRampPalette(colors = c("#ffffff", "#6666ff"))(100), zlim = c(0,1),
    legend.shrink = .5, legend.lab = "Confidence", legend.line = 3, legend.cex = 1)
  title(ylab = "Relative cum. time spent", line = 3, cex.lab = 1.5)
  grid()
  curve(dev_p3, 0, 1, col = "blue", add = TRUE)
  
  curve(dev_ci_upper_p3avg, 0, 1, col = "#0000ff33", add = TRUE, lty = 2)
  curve(dev_ci_lower_p3avg, 0, 1, col = "#0000ff33", add = TRUE, lty = 3)
  
  for (i in 1:length(all_signals)) {
    pId <- names(all_signals)[i]
    tempf <- all_signals[[pId]]$DEV$get0Function()
    curve(tempf, 0, 1, col="#00000033", lty=if (omega[i] == 0) 3 else 2, add=TRUE)
  }
  
  legend(0.01, .98, legend = c("DEV (avg.)"),
         col = "blue", lty = 1, lwd = 2.5, cex = 1.3)
  
  
  tempz <- fields::interp.surface.grid(obj = z_desc, grid.list = list(
    x = seq(from = 0, to = 1, length.out = 90), y = seq(from = 0, to = 1, length.out = 90)))
  fields::image.plot(x = tempz$x, y = tempz$y, z = tempz$z,
    xlab = "", ylab = "",
    col = colorRampPalette(colors = c("#ffffff", "forestgreen"))(100),
    ylim = c(0, desc_ci_upper_p3avg(1) * 1.1), zlim = c(0, 1),
    legend.shrink = .5, legend.lab = "Confidence", legend.line = 3, legend.cex = 1)
  title(xlab = "Relative project time", ylab = "Relative cum. time spent", line = 3, cex.lab = 1.5)
  grid()
  curve(desc_p3, 0, 1, col = "forestgreen", add = TRUE)
  
  curve(desc_ci_upper_p3avg, 0, 1, col = "#228B2233", add = TRUE, lty = 2)
  curve(desc_ci_lower_p3avg, 0, 1, col = "#228B2233", add = TRUE, lty = 3)
  
  for (i in 1:length(all_signals)) {
    pId <- names(all_signals)[i]
    tempf <- all_signals[[pId]]$DESC$get0Function()
    curve(tempf, 0, 1, col="#00000033", lty=if (omega[i] == 0) 3 else 2, add=TRUE)
  }
  
  legend(0.01, 0.98 * desc_ci_upper_p3avg(1) * 1.1, legend = c("DESC (avg.)"),
         col = "forestgreen", lty = 1, lwd = 2.5, cex = 1.3)
})

eval(plotexpr)
```
```{r echo=FALSE, eval=FALSE}
# saveTikzExpr(name = "emp_ci_req_dev", height = 4.2, expr = plotexpr)
# saveEPSExpr(name = "emp_ci_req_dev", height = 4.2, expr = plotexpr)
pdf(file = "../figures/fig_fdfe_emp_ci_req_dev.pdf", width = fe_fig_width, height = 5.9, pointsize = 12)
#png(filename = "../figures/fig_fdfe_emp_ci_req_dev.png", width = 2000, height = 2000 / fe_fig_width * 4.2, pointsize = 40)
eval(plotexpr)
dev.off()
```

```{r echo=FALSE, eval=FALSE}
saveTikzExpr(name = "emp_ci_req_dev", height = 7.7, expr = plotexpr)
```


```{r}
z_a <- loadResultsOrCompute(file = "../results/ci_p3avg_z_a.rds", computeExpr = {
  compute_z_p3avg(varname = "A", x = x, y = y)
})
z_cp <- loadResultsOrCompute(file = "../results/ci_p3avg_z_cp.rds", computeExpr = {
  compute_z_p3avg(varname = "CP", x = x, y = y)
})
z_freq <- loadResultsOrCompute(file = "../results/ci_p3avg_z_freq.rds", computeExpr = {
  compute_z_p3avg(varname = "FREQ", x = x, y = y)
})
z_scd <- loadResultsOrCompute(file = "../results/ci_p3avg_z_scd.rds", computeExpr = {
  compute_z_p3avg(varname = "SCD", x = x, y = y)
})


plotexpr <- rlang::expr({
  par(mfrow=c(4,1), mar=c(2,4,1,.8), family = "LM Roman 10", cex.axis = 1.5, oma=rep(0,4))
  # par(mar = c(5,4,0,1), family = "LM Roman 10", cex.axis = 1.1)
  # layout(mat = matrix(data = c(1,0,2, 0,0,0, 3,0,4), byrow = TRUE, ncol = 3),
  #        heights = c(10,2,10), widths = c(10,2,10))
  
  
  
  tempz <- fields::interp.surface.grid(obj = z_a, grid.list = list(
    x = seq(from = 0, to = 1, length.out = 90), y = seq(from = 0, to = 1, length.out = 90)))
  fields::image.plot(x = tempz$x, y = tempz$y, z = tempz$z,
    xlab = "", ylab = "",
    col = colorRampPalette(colors = c("#ffffff", "#f8766d"))(100), zlim = c(0,1),
    legend.shrink = .5, legend.lab = "Confidence", legend.line = 3)#, legend.cex = 1)
  title(ylab = "Relative amount of activity", line = 2.5, cex.lab = 1.2)
  grid()
  
  curve2(p3_avg_signals_all$A$get0Function(), 0, 1, col = "red", add = TRUE)
  curve2(a_ci_upper_p3avg, 0, 1, col = "#f8766d33", add = TRUE, lty = 2)
  curve2(a_ci_lower_p3avg, 0, 1, col = "#f8766d33", add = TRUE, lty = 3)
  
  temp <- append(project_signals, project_signals_2nd_batch)
  for (i in 1:length(temp)) {
    pId <- names(temp)[i]
    tempf <- temp[[pId]]$A$get0Function()
    curve2(tempf, 0, 1, col="#00000033", lty=if (omega[i] == 0) 3 else 2, add=TRUE)
  }
  
  legend(0.01, .98, legend = c("A (avg.)", "Weight $=0$", "Weight $>0$"),# bg = "white",
         col = c("red", rep("#00000033", 2)), lty = c(1, 3, 2), lwd = c(2.5, 2, 2), cex = 1.2)
  
  
  
  
  tempz <- fields::interp.surface.grid(obj = z_cp, grid.list = list(
    x = seq(from = 0, to = 1, length.out = 90), y = seq(from = 0, to = 1, length.out = 90)))
  fields::image.plot(x = tempz$x, y = tempz$y, z = tempz$z,
    xlab = "", ylab = "",
    col = colorRampPalette(colors = c("#ffffff", "#7CAE00"))(100), zlim = c(0,1),
    legend.shrink = .5, legend.lab = "Confidence", legend.line = 3, legend.cex = 1)
  title(ylab = "Relative amount of activity", line = 2.5, cex.lab = 1.2)
  grid()
  
  curve2(p3_avg_signals_all$CP$get0Function(), 0, 1, col = "darkgreen", add = TRUE)
  curve2(cp_ci_upper_p3avg, 0, 1, col = "#0000ff33", add = TRUE, lty = 2)
  curve2(cp_ci_lower_p3avg, 0, 1, col = "#0000ff33", add = TRUE, lty = 3)
  
  temp <- append(project_signals, project_signals_2nd_batch)
  for (i in 1:length(temp)) {
    pId <- names(temp)[i]
    tempf <- temp[[pId]]$CP$get0Function()
    curve2(tempf, 0, 1, col="#00000033", lty=if (omega[i] == 0) 3 else 2, add=TRUE)
  }
  
  legend(0.01, .98, legend = c("CP (avg.)"),
         col = "darkgreen", lty = 1, lwd = 2.5, cex = 1.2)
  
  
  
  
  tempz <- fields::interp.surface.grid(obj = z_freq, grid.list = list(
    x = seq(from = 0, to = 1, length.out = 90), y = seq(from = 0, to = 1, length.out = 90)))
  fields::image.plot(x = tempz$x, y = tempz$y, z = tempz$z,
    xlab = "", ylab = "",
    col = colorRampPalette(colors = c("#ffffff", "#00BFC4"))(100), zlim = c(0,1),
    legend.shrink = .5, legend.lab = "Confidence", legend.line = 3, legend.cex = 1)
  title(ylab = "Relative amount of activity", line = 2.5, cex.lab = 1.2)
  grid()
  
  curve2(p3_avg_signals_all$FREQ$get0Function(), 0, 1, col = "blue", add = TRUE)
  curve2(freq_ci_upper_p3avg, 0, 1, col = "#0000ff33", add = TRUE, lty = 2)
  curve2(freq_ci_lower_p3avg, 0, 1, col = "#0000ff33", add = TRUE, lty = 3)
  
  temp <- append(project_signals, project_signals_2nd_batch)
  for (i in 1:length(temp)) {
    pId <- names(temp)[i]
    tempf <- temp[[pId]]$FREQ$get0Function()
    curve2(tempf, 0, 1, col="#00000033", lty=if (omega[i] == 0) 3 else 2, add=TRUE)
  }
  
  legend(0.01, .98, legend = c("FREQ (avg.)"),
         col = "blue", lty = 1, lwd = 2.5, cex = 1.2)
  
  
  
  
  tempz <- fields::interp.surface.grid(obj = z_scd, grid.list = list(
    x = seq(from = 0, to = 1, length.out = 90), y = seq(from = 0, to = 1, length.out = 90)))
  fields::image.plot(x = tempz$x, y = tempz$y, z = tempz$z,
    xlab = "", ylab = "",
    col = colorRampPalette(colors = c("#ffffff", "#C77CFF"))(100), zlim = c(0,1),
    legend.shrink = .5, legend.lab = "Confidence", legend.line = 3, legend.cex = 1)
  title(ylab = "Source code density", line = 2.5, cex.lab = 1.2)
  grid()
  
  curve2(p3_avg_signals_all$SCD$get0Function(), 0, 1, col = "purple", add = TRUE)
  curve2(scd_ci_upper_p3avg, 0, 1, col = "#0000ff33", add = TRUE, lty = 2)
  curve2(scd_ci_lower_p3avg, 0, 1, col = "#0000ff33", add = TRUE, lty = 3)
  
  temp <- append(project_signals, project_signals_2nd_batch)
  for (i in 1:length(temp)) {
    pId <- names(temp)[i]
    tempf <- temp[[pId]]$SCD$get0Function()
    curve2(tempf, 0, 1, col="#00000033", lty=if (omega[i] == 0) 3 else 2, add=TRUE)
  }
  
  legend(0.325, .2, legend = c("SCD (avg.)"),
         col = "purple", lty = 1, lwd = 2.5, cex = 1.2)
})

eval(plotexpr)
```

```{r}
pdf(file = "../figures/fig_fdfe_emp_ci_req_dev_sc.pdf", width = 3.33, height = 7.5, pointsize = 9)
eval(plotexpr)
dev.off()
```


```{r}
saveTikzExpr(name = "emp_ci_req_dev_sc", height = 7, width = 3.525, expr = plotexpr)
```




```{r echo=FALSE, eval=FALSE}
all_projects_acp <- ggplot(data = student_projects, aes(x = length(label), fill = label)) +
  geom_bar() + facet_grid(label ~ project, labeller = ggplot2::labeller(
    label = `names<-`(c("Adaptive", "Corrective", "Perfective"), c("a", "c", "p")),
    project = `names<-`(paste0("Pr. ", 1:9), paste0("project_", 1:9))
  )) +
  fe_gg2_style(facets = TRUE) +
  theme(legend.position = "none", axis.text.x.bottom = element_blank(), axis.title.x.bottom = element_blank(), axis.title.y.left = element_blank()) +
  #scale_fill_manual(values = c("red", "orange", )) +
  scale_fill_grey()

saveTikzAndEps(ggInst = all_projects_acp, name = "all_projects_acp", height = 3)
```



```{r echo=FALSE, eval=FALSE}
temp <- doWithParallelCluster(expr = {
  g <- expand.grid(data.frame(
    x = seq(0.4, 0.9, length.out=20),
    y = seq(0.4, 0.8, length.out=20)
  ))
  
  foreach::foreach(
    idx = rownames(g),
    .inorder = FALSE,
    .combine = rbind
  ) %dopar% {
    p <- g[idx,]
    
    matrix(data = arrow_dir_smooth(x = p$x, y = p$y), nrow = 1)
  }
})
```


```{r echo=FALSE, eval=FALSE}
plotexpr <- expression({
  temp <- temp[, 1:4]
  temp <- cbind(temp, sqrt(temp[, 3]^2 * temp[, 4]^2))
  par(mfrow=c(1,1), mar=c(4,4,.8,.8), family = "LM Roman 10")#, oma=rep(0,4))
  
  fields::image.plot(x = z_dev$x, y = z_dev$y, z = z_dev$z,
    xlab = "", ylab = "",
    col = colorRampPalette(colors = c("#ffffff", "#6666ff"))(100),
    zlim = c(0,1), # make colors weaker
    xlim = c(0.4,0.9), ylim = c(0.4,0.6),
    legend.shrink = .5, legend.lab = "Confidence", legend.line = 2.5)
  title(xlab = "Relative project time", ylab = "Relative cum. time spent", line = 2.5)
  grid()
  
  
  curve(dev_ci_upper_p3avg, 0, 1, add = TRUE, col = "blue", lty = 2)
  curve(dev_ci_lower_p3avg, 0, 1, add = TRUE, col = "blue", lty = 2)
  # tempf <- function(x) all_signals$Project5$DEV$get0Function()(x)
  # curve(tempf, 0, 1, add = TRUE, col = "red", lwd = 2)
  
  for (i in 1:length(all_signals)) {
    pId <- names(all_signals)[i]
    tempf <- all_signals[[pId]]$DEV$get0Function()
    curve(tempf, 0, 1, col="#00000033", lty=if (omega[i] == 0) 3 else 2, add=TRUE)
  }
  
  for (i in 1:nrow(temp)) {
    r <- temp[i, ]
    to <- r[3:4]
    if (r[2] < dev_ci_lower_p3avg(r[1]) || r[2] > dev_ci_upper_p3avg(r[1])) {
      next
    }
    #to <- to / max(abs(to)) * 0.01
    to <- to / max(abs(to)) * 0.0175 * (.6 + .4 * log(r[5]) / log(max(temp[, 5])))
    arrows(x0 = r[1] + (if (to[1] < 0) abs(to[1]) else 0), y0 = r[2] + (if (to[2] < 0) abs(to[2]) else 0),
           x1 = r[1] + (if (to[1] < 0) abs(to[1]) else 0) + to[1], y1 = r[2] + (if (to[2] < 0) abs(to[2]) else 0) + to[2], length = 0.05, lwd = .8, col = "#333377")
  }
  
  
})

eval(plotexpr)
```

```{r echo=FALSE, eval=FALSE}
saveTikzExpr(expr = plotexpr, name = "vectorfield_dev", height = 2.8)
saveEPSExpr(expr = plotexpr, name = "vectorfield_dev", height = 2.8)
```

```{r echo=FALSE, eval=FALSE}
pdf(file = "../figures/fig_fdfe_vectorfield_dev.pdf", width = fe_fig_width, height = 2.3, pointsize = 8)
eval(plotexpr)
dev.off()
```


```{r}
plotexpr <- expression({
  par(mfrow=c(2,2), mar=c(4,4,.8,0), family = "LM Roman 10")
  use_func <- all_signals$Project4$REQ$get0Function()
  gbar <- req_p3
  
  tempz <- fields::interp.surface.grid(obj = z_req, grid.list = list(
    x = seq(from = 0, to = 1, length.out = 80), y = seq(from = 0, to = 1, length.out = 80)))
  
  image(x = tempz$x, y = tempz$y, z = tempz$z,
    xlab = "", ylab = "",
    col = colorRampPalette(colors = c("#ffffff", "#ff6666"))(100), zlim = c(0,1))
  grid()
  title(xlab = "Relative project time", ylab = "Rel. cum. time spent", line = 2.5)
  curve(gbar, 0, 1, col = "#666666", lwd = 1.5, xlab = "", ylab = "", add = TRUE)
  curve(use_func, 0, 1, col = "red", lwd = 2, add = TRUE)
  curve(req_ci_upper_p3avg, 0, 1, col = "#bb000066", lwd = 1.5, lty = 3, add = TRUE)
  curve(req_ci_lower_p3avg, 0, 1, col = "#bb000066", lwd = 1.5, lty = 2, add = TRUE)
  
  legend(0.01, 0.98, legend = c("Avg.", "Pr. 4"), col = c("#666666", "red"), lwd = c(1.5,2), cex = .9, bg = "white")
  legend(0.44, 0.4, legend = c("Upper CI", "Lower CI"), col = "red", lty = c(3,2), lwd = 1.5, bg = "white")
  
  
  
  tempfo <- function(x) {
    (.3 * tempf(x) + .7 * (1 - tempf1(x)))
    # sapply(X = x, FUN = function(x_) {
    #   (CI_req_p3avg(x = x_, y = use_func(x_)) + 1 - abs(req_p3(x_) - min(req_ci_upper_p3avg(x_), max(req_ci_lower_p3avg(x_), use_func(x_))))) / 2
    # })
  }
  curve(tempfo, 0, 1, ylim = c(0,1), col = "Maroon", xlab = "", ylab = "", lwd = 2)
  grid()
  title(xlab = "Relative project time", ylab = "Objective of $\\mathsf{M}$\\textsuperscript{conf} and $\\mathsf{M}$\\textsuperscript{dist}", line = 2.5)
  legend(0.01, 0.4, legend = c("Objective using", "$\\rho=0.3$ and $\\sigma=0.7$"), col = c("Maroon", "white"), lty = 1, lwd = 2, bg = "white")
  # curve(tempf, add = TRUE, col="red")
  # curve(tempf1, add = TRUE, col="blue")
  
  
  
  tempf <- function(x) {
    sapply(X = x, FUN = function(x_) {
      CI_req_p3avg(x = x_, y = use_func(x_))
    })
  }
  
  curve(tempf, 0, 1, xlab = "", ylab = "", col = "DarkCyan", ylim = c(0, 1), lwd = 2)
  grid()
  title(xlab = "Relative project time", ylab = "Empirical confidence", line = 2.5)
  legend(0.01, 1, legend = c("Confidence"), col = "DarkCyan", lty = 1, lwd = 2, bg = "white")
  
  
  
  tempf1 <- function(x) {
    sapply(X = x, FUN = function(x_) {
      abs(req_p3(x_) - min(req_ci_upper_p3avg(x_), max(req_ci_lower_p3avg(x_), use_func(x_))))
    })
  }
  curve(tempf1, 0, 1, col = "DarkOrange", xlab = "", ylab = "", ylim = c(0,0.45), lwd = 2)
  grid()
  title(xlab = "Relative project time", ylab = "Distance to avg. REQ", line = 2.5)
  
  tempf2 <- function(x) {
    sapply(X = x, FUN = function(x_) {
      max(req_p3(x_) - req_ci_lower_p3avg(x_), req_ci_upper_p3avg(x_) - req_p3(x_))
    })
  }
  curve(tempf2, 0, 1, col = "DarkOrange", add = TRUE, lty = 2, lwd = 1.5)
  legend(0.25, 0.45, legend = c("Distance", "Distance (max)"), col = "DarkOrange", lty = c(1,2), lwd = c(2,1.5), bg = "white")
})
eval(plotexpr)
```

```{r}
cubature::cubintegrate(f = tempfo, 0, 1)$integral
```


```{r}
readRDS(file = "../results/ci_p3avg_Lavg-test.rds")
readRDS(file = "../results/ci_p3avg_Larea-test.rds")
```
```{r}
#saveEPSExpr(expr = plotexpr, name = "empconf_areadist_req", height = 3.6)
saveTikzExpr(expr = plotexpr, name = "empconf_areadist_req", height = 3.6)
```
```{r}
cairo_ps("../figures/fig_fdfe_empconf_areadist_req.eps", family = "LM Roman 10", width = fe_fig_width, height = 3.6)
eval(plotexpr)
dev.off()
```
```{r}
pdf(file = "../figures/fig_fdfe_empconf_areadist_req.pdf", width = fe_fig_width, height = 3.6, pointsize = 10, family = "LM Roman 10")
eval(plotexpr)
dev.off()
```






```{r echo=FALSE, eval=FALSE}
temp <- doWithParallelCluster(expr = {
  g <- expand.grid(data.frame(
    x = seq(0.45, 0.8, length.out=12),
    y = seq(0.3, .95, length.out=12)
  ))
  
  foreach::foreach(
    idx = rownames(g),
    .inorder = FALSE,
    .combine = rbind
  ) %dopar% {
    p <- g[idx,]
    
    matrix(data = arrow_dir_smooth(x = p$x, y = p$y), nrow = 1)
  }
})
```



```{r echo=FALSE, eval=FALSE}
plotexpr <- expression({
  temp <- temp[, 1:4]
  temp <- cbind(temp, sqrt(temp[, 3]^2 * temp[, 4]^2))
  par(mfrow=c(1,1), mar=c(4,4,.8,.8), family = "LM Roman 10")
  
  tempz <- fields::interp.surface.grid(obj = z_dev, grid.list = list(
    x = seq(from = 0, to = 1, length.out = 125), y = seq(from = 0, to = 1, length.out = 125)))
  
  # plot(x=0, y=0, xlab = "", ylab = "", xlim = c(0.55,0.8), ylim = c(0.4,0.9))
  fields::image.plot(x = tempz$x, y = tempz$y, z = tempz$z,
    xlab = "", ylab = "",
    col = colorRampPalette(colors = c("#ffffff", "#6666ff"))(100),
    zlim = c(0,1), # make colors weaker
    xlim = c(0.45,0.8), ylim = c(0.3,.95),
    legend.shrink = .5, legend.lab = "Confidence", legend.line = 2.5)
  title(xlab = "Relative project time", ylab = "Relative cum. time spent", line = 2.5)
  grid()
  
  
  # curve(dev_ci_upper_p3avg, 0, 1, add = TRUE, col = "blue", lty = 2)
  # curve(dev_ci_lower_p3avg, 0, 1, add = TRUE, col = "blue", lty = 2)
  # tempf <- function(x) all_signals$Project5$DEV$get0Function()(x)
  # curve(tempf, 0, 1, add = TRUE, col = "red", lwd = 2)
  
  # for (i in 1:length(all_signals)) {
  #   pId <- names(all_signals)[i]
  #   tempf <- all_signals[[pId]]$DEV$get0Function()
  #   curve(tempf, 0, 1, col="#00000033", lty=if (omega[i] == 0) 3 else 2, add=TRUE)
  # }
  
  for (i in 1:nrow(temp)) {
    r <- temp[i, ]
    to <- r[3:4]
    if (r[2] < dev_ci_lower_p3avg(r[1]) || r[2] > dev_ci_upper_p3avg(r[1])) {
      next
    }
    #to <- to / max(abs(to)) * 0.01
    to <- to / max(abs(to)) * 0.0175 * (.6 + .4 * log(r[5]) / log(max(temp[, 5])))
    arrows(x0 = r[1] + (if (to[1] < 0) abs(to[1]) else 0), y0 = r[2] + (if (to[2] < 0) abs(to[2]) else 0),
           x1 = r[1] + (if (to[1] < 0) abs(to[1]) else 0) + to[1], y1 = r[2] + (if (to[2] < 0) abs(to[2]) else 0) + to[2], length = 0.05, lwd = 1.5, col = "#333377")
  }
  
  polygon(
    x = c(fch_x, rev(fch_x)),
    y = c(fch_80_upper(fch_x), fch_95_upper(rev(fch_x))),
    border = NA,
    col = "#00000011"
  )
  
  polygon(
    x = c(fch_x, rev(fch_x)),
    y = c(fch_80_lower(fch_x), fch_95_lower(rev(fch_x))),
    border = NA,
    col = "#00000011"
  )
  
  polygon(
    x = c(fch_x, rev(fch_x)),
    y = c(fch_80_upper(fch_x), fch_80_lower(rev(fch_x))),
    border = NA,
    col = "#00000033"
  )
  
  curve(dev_ci_lower_p3avg, 0, 1, col = "#0000ff66", add = TRUE, lty = 2)
  rect(xleft = .6625, ybottom = .325, xright = .7375, ytop = .45, col = "white")
  segments(x0 = .6625, y0 = .325, x1 = .65, y1 = fch_mean(.65), col = "#666666")
  segments(x0 = .6625, y0 = .45, x1 = .65, y1 = fch_mean(.65), col = "#666666")
  segments(x0 = .7375, y0 = .325, x1 = .75, y1 = fch_95_lower(.75), col = "#666666")
  segments(x0 = .7375, y0 = .45, x1 = .75, y1 = fch_95_upper(.75), col = "#666666")
  arrows(x0 = .675, y0 = .41, x1 = .725, y1 = .41 + (.725-.675)/ed_tc_x*ed_tc_y, lwd = 2, length = .125, col = "#800080")
  
  curve(fch_80_upper, 0.65, 0.75, col="lightgrey", add = TRUE)
  curve(fch_80_lower, 0.65, 0.75, col="lightgrey", add = TRUE)
  curve(fch_95_upper, 0.65, 0.75, col="darkgrey", add = TRUE)
  curve(fch_95_lower, 0.65, 0.75, col="darkgrey", add = TRUE)
  
  curve(dev_p3, 0, 1, col = "blue", lwd = 2, add = TRUE)
  curve(dev_ci_upper_p3avg, 0, 1, col = "#0000ff66", add = TRUE, lty = 3)
  
  # segments(x0 = .625, y0 = .93, x1 = .65, y1 = fch_mean(.65))
  # segments(x0 = .625, y0 = .805, x1 = .65, y1 = fch_mean(.65))
  # segments(x0 = .7, y0 = .93, x1 = .75, y1 = fch_95_upper(.75))
  # segments(x0 = .7, y0 = .805, x1 = .75, y1 = fch_95_lower(.75))
  # rect(xleft = .625, ybottom = .805, xright = .7, ytop = .93, col = "white")
  # arrows(x0 = .675, y0 = .415, x1 = .725, y1 = .415 + (.725-.675)/ed_tc_x*ed_tc_y, lwd = 2, length = .2, col = "#800080")
  
  curve(pr5_dev, 0, 0.65, col = "maroon", lty = 1, lwd = 2, add = TRUE)
  curve(pr5_dev, 0.65, 0.75, col = "maroon", lty = 4, lwd = 2, add = TRUE)
  curve(fch_mean, 0.65, 0.75, col = "red", lty = 2, lwd = 2, add = TRUE)
  #arrows(x0 = 0.65, x1 = 0.75, y0 = 0.65, y1 = 0.65 + 0.1 / ed_tc_x * ed_tc_y, lwd = 2, length = .2, col = "#800080")
  
  legend(0.455, .93, legend = c("Avg. DEV", "Upper CI", "Lower CI", "Pr. 5 DEV", "Pr. 5 actual", "Pr. 5 prediction", "Avg. direction"),
         bg="white", col = c(rep("blue", 3), rep("maroon", 2), "red", "#800080"), cex = .85,
         lty = c(1, 3, 2, 1, 4, 2, 1),
         lwd = c(2, 1.5, 1.5, 2, 1.5, 2, 2, 2))
})

eval(plotexpr)
```

```{r}
#saveEPSExpr(expr = plotexpr, name = "forecast_vectorfield", height = 2.3)
saveTikzExpr(expr = plotexpr, name = "forecast_vectorfield", height = 3.2)
```


```{r}
plotexpr <- expression({
  par(mfrow=c(1,1), mar=c(4,4,.8,1), family = "LM Roman 10")
  
  align_srbtaw_rss_surface <- readRDS(file = "../results/align_srbtaw_rss_surface.rds")
  z_temp <- list(
    x = seq(0, 1, by = minStep),
    y = seq(0, 1, by = minStep),
    z = align_srbtaw_rss_surface)
  
  fields::image.plot(z_temp, col = hcl.colors(101, "YlOrRd", rev = TRUE),
                     legend.shrink = .5, legend.lab = "Loss ($\\log$(RSS))", legend.line = 2.5,
                     xlab = "", ylab = "")
  title(ylab = "$t$\\textsubscript{begin}", xlab = "$t$\\textsubscript{end}", line = 2.5)
  grid()
  
  z_temp$x <- z_temp$x[2:51]
  z_temp$y <- z_temp$y[2:51]
  contour(z_temp, col = "#222222", add = TRUE)
  
  temp <- align_srbtaw_rss_surface
  temp[is.na(temp)] <- Inf
  temp <- which(temp == min(temp), arr.ind = TRUE)
  points(x = temp[, 1] / use_resol, y = temp[, 2] / use_resol, pch = 2, col = "red", cex = 2)
  
  legend(.01, .98, legend = c("Lowest loss ($\\approx5.672$)"),
         pch = 2, col = "red", bg = "white", cex = 1)
})

eval(plotexpr)
```

```{r}
saveTikzExpr(expr = plotexpr, name = "srbtaw_rss_loss", height = 2.4)
```


```{r}
plotexpr <- expression({
  temp <- t(pAll_corr[1:5,]) # skip last pattern (3b[1-16]) for the paper
  rownames(temp) <- c(rownames(temp)[1:24], "ImpFac_m", "ImpFac_p")
  colnames(temp) <- c("I", "II", "(*) II", "III", "(*) III")
  temp[is.na(temp)] <- 0
  #rownames(temp) <- gsub("_", "\\\\_", rownames(temp))
  rownames(temp) <- sapply(X = rownames(temp), function(n) {
    sp <- strsplit(x = n, split = "_")[[1]]
    paste0("$\\mathsf{", sp[1], "}$", "\\textsubscript{", sp[2], "}")
  })
  temp <- temp[, rev(colnames(temp))]
  
  ggcorrplot::ggcorrplot(corr = temp, ggtheme = fe_gg2_style() + theme(
      axis.text.x.top = element_text(angle = 90, vjust = 1/3, hjust = 0, size = 9),
      #axis.text.y.left = element_text(size = 9),
      legend.position = "bottom",
      legend.text = element_text(vjust = 0),
      legend.title = element_text(margin = margin(r = 10, b = 3)),
      legend.key.width = unit(20, 'pt'),
      legend.key.height = unit(8, 'pt'),
      axis.title.y.left = element_blank()
    )
  ) +
  scale_x_discrete(position = "top") +
  #scale_y_discrete(limits = rev) +
  scale_fill_gradient2(
    low = "#0037AA", mid = "#eeeeee", high = "#AA0037", guide = "colorbar",
    limits = c(-1,1), breaks = seq(-1, 1, by = .5)) +
  labs(fill = "Pearson sample correlation")
})

eval(plotexpr)
```

```{r}
saveTikzAndEps(ggInst = eval(plotexpr), name = "detect-sc-corr", height = 2.3)
```



# Correlations of scores for issue-tracking data

We want to produce one large, common correlation-table (not a matrix) using process models 1-3 and 4(1-3).

```{r message=FALSE}
plotexpr <- expression({
  temp <- t(rbind(
    `rownames<-`(x = pAll_it_corr, value = c("I", "II", "III")),
    `rownames<-`(x = p4All_it_corr, value = c("*I", "*II", "*III"))
  ))
  temp <- temp[, rev(colnames(temp))]
  
  rownames(temp) <- sapply(X = rownames(temp), function(n) {
    sp <- strsplit(x = n, split = "_")[[1]]
    paste0("$\\mathsf{", sp[1], "}_{\\mathrm{", sp[2], "}}$")
  })
  
  
  ggcorrplot::ggcorrplot(corr = temp, ggtheme = fe_gg2_style() + theme(
      axis.text.x.top = element_text(angle = 90, vjust = 1/3, hjust = 0, size = 8),
      axis.text.y.left = element_text(size = 8),
      legend.position = "bottom",
      legend.text = element_text(vjust = 0),
      legend.title = element_text(margin = margin(r = 10, b = 3)),
      legend.key.width = unit(20, 'pt'),
      legend.key.height = unit(8, 'pt'),
      axis.title.y.left = element_blank())) +
  scale_x_discrete(position = "top") +
  scale_fill_gradient2(
    low = "#0037AA", mid = "#eeeeee", high = "#AA0037", guide = "colorbar",
    limits = c(-1,1), breaks = seq(-1, 1, by = .5)) +
  labs(fill = "Pearson sample correlation")
})

eval(plotexpr)
```

```{r message=FALSE}
saveTikzAndEps(ggInst = eval(plotexpr), name = "detect-it-corr", height = 4)
```
```{r}
tikzDevice::tikz(file = paste0("temp", name, ".tex"), width = fe_fig_width, verbose = FALSE)
eval(plotexpr)
dev.off()
```



















