---
title: "Reproduction Material for article ``Machine-aided operationalization of continuous software processes''"
author: "Sebastian Hönel"
date: "`r format(Sys.time(), '%B %d, %Y')`"
bibliography: ../inst/REFERENCES.bib
urlcolor: blue
output:
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 6
    df_print: kable
    keep_tex: yes
  md_document:
    toc: yes
    toc_depth: 6
    df_print: kable
    variant: gfm
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 6
    toc_float: yes
    df_print: kable
  word_document: default
header-includes:
- \usepackage{bm}
- \usepackage{mathtools}
- \usepackage{xurl}
---

\newcommand*\mean[1]{\overline{#1}}
\newcommand{\norm}[1]{\left\lvert#1\right\rvert}
\newcommand{\infdiv}[2]{#1\;\|\;#2}

```{r setoptions, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
opts_chunk$set(tidy = TRUE, tidy.opts = list(indent=2))
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
source(file = "../helpers.R")
source(file = "./common-funcs.R")
source(file = "../models/modelsR6.R")
source(file = "../models/SRBTW-R6.R")

library(ggplot2)
library(ggpubr)
library(formatR)
```



# Overview

This notebook represents the entire reproduction material required to reproduce and replicate the presented results, tables, and figures as used in the paper "Machine-aided operationalization of continuous software processes".

Please refer to the full technical report [@honel2021technical] for all technical details, code, and data.


```{r}
if (!endsWith(getwd(), "notebooks")) {
  stop("Must be in the notebooks working directory.")
}
```


# Research Questions

```{r}
ground_truth <- read.csv(file = "../data/ground-truth.csv", sep = ";")
ground_truth$consensus_score <- ground_truth$consensus / 10
```

## RQ1(a)

Here, we will produce some plots that show the correlation between computed scores.

```{r}
p3_it_scores <- loadResultsOrCompute(file = "../results/p3_it_scores.rds", computeExpr = {
  as.data.frame(
    compute_all_scores_it(
      alignment = p3_it_projects, patternName = "p3_it", vartypes = names(p3_it_signals)))
})
```


```{r fig.height=10, fig.width=10}
temp <- stats::cor(cbind(data.frame(gt_consensus = ground_truth$consensus_score), p3_it_scores[,]))
temp[is.na(temp)] <- 0 # The scores for 'DEV_PEAK' are all 1 and the correlation therefore is NA

corrplot::corrplot(corr = temp, type = "upper",# order = "hclust",
                   tl.col = "black", tl.srt = 90)
```

For issue-tracking data, we have the correlation of scores on a per-variable basis, but not for source code. So next, we will aggregate the scores using the mean across scores. For example, we will produce the average of all `area`-scores.

```{r}
scores <- unique(gsub(pattern = "^(REQ|DEV|DESC)_", replacement = "", x = colnames(p3_it_scores)))
# Remove mean absolute error (mae), as it's the same as area in our case, since the segment length equaled 1.
scores <- scores[!(scores %in% "mae")]
scores
```

```{r}
p3_it_scores_mean <- data.frame(row.names = rownames(p3_it_scores))

for (name in scores) {
  temp <- cbind(p3_it_scores[, paste0("REQ_", name)], p3_it_scores[, paste0("DEV_", name)], p3_it_scores[, paste0("DESC_", name)])
  temp.mean <- as.vector(apply(X = temp, MARGIN = 1, FUN = mean))
  p3_it_scores_mean <- cbind(p3_it_scores_mean, `colnames<-`(matrix(data = temp.mean, ncol = 1), name))
}

p3_it_scores_mean
```

Now with this data, we can show a correlation plot:

```{r}
plotexpr <- expression({
  temp <- stats::cor(cbind(data.frame(GroundTruth = ground_truth$consensus_score), p3_it_scores_mean[,]))
  temp[is.na(temp)] <- 0 # The scores for 'DEV_PEAK' are all 1 and the correlation therefore is NA
  corrplot::corrplot(corr = temp, type = "upper", tl.col = "black", tl.srt = 90)
})

eval(plotexpr)
```

```{r}
png(filename = "../figures/reproduction/rq1_corr_it.png", width = 400, height = 400)
eval(plotexpr)
invisible(dev.off())
```


Let's load the scores for Source code. Note that these were previously aggregated once using the mean (suffix `_m`) and once using the product (`_p`) across __all__ variables. We will keep the mean for now, and then rename the columns.

```{r}
# P3("avg, no align"actually)
p3_sc_scores <- loadResultsOrCompute(file = "../results/p3_avg_no_scores.rds", computeExpr = {
  as.data.frame(compute_all_scores(alignment = p3_avg_no_align, patternName = "p3_avg"))
})

temp <- NULL
for (name in scores) {
  temp.scores <- `colnames<-`(data.frame(x = p3_sc_scores[, paste0(name, "_m")]), name)
  temp <- if (is.null(temp)) temp.scores else cbind(temp, temp.scores)
}

p3_sc_scores <- temp
p3_sc_scores
```

```{r}
plotexpr <- expression({
  temp <- stats::cor(cbind(data.frame("GroundTruth" = ground_truth$consensus_score), p3_sc_scores[,]))
  temp[is.na(temp)] <- 0 # The scores for 'DEV_PEAK' are all 1 and the correlation therefore is NA
  
  corrplot::corrplot(corr = temp, type = "upper", tl.col = "black", tl.srt = 90)
})

eval(plotexpr)
```

```{r}
png(filename = "../figures/reproduction/rq1_corr_sc.png", width = 400, height = 400)
eval(plotexpr)
invisible(dev.off())
```


Since obviously, the metrics have a different correlation between the two types of data, we should compute the pairwise correlations between each score of source code and issue-tracking:

```{r}
temp <- matrix(nrow = 1, ncol = 0)
for (name in scores) {
  temp <- cbind(temp, stats::cor(p3_it_scores_mean[, name], p3_sc_scores[, name]))
}
colnames(temp) <- scores
temp
```

Next, we can examine the correlations between activities. We will do this for issue-tracking data.

```{r}
p3_it_scores
```




# References {-}

<div id="refs"></div>

