---
title: "Reproduction Material for article ``Machine-aided operationalization of continuous software processes''"
author: "Sebastian Hönel"
date: "`r format(Sys.time(), '%B %d, %Y')`"
bibliography: ../inst/REFERENCES.bib
urlcolor: blue
output:
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 6
    df_print: kable
    keep_tex: yes
  md_document:
    toc: yes
    toc_depth: 6
    df_print: kable
    variant: gfm
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 6
    toc_float: yes
    df_print: kable
  word_document: default
header-includes:
- \usepackage{bm}
- \usepackage{mathtools}
- \usepackage{xurl}
- \usepackage[nottoc]{tocbibind}
---

\newcommand*\mean[1]{\overline{#1}}
\newcommand{\norm}[1]{\left\lvert#1\right\rvert}
\newcommand{\infdiv}[2]{#1\;\|\;#2}


\listoffigures

\listoftables



```{r setoptions, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
opts_chunk$set(tidy = TRUE, tidy.opts = list(indent=2))
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
source(file = "../helpers.R")
source(file = "./common-funcs.R")
source(file = "../models/modelsR6.R")
source(file = "../models/SRBTW-R6.R")

library(ggplot2)
library(ggpubr)
library(formatR)
```



# Overview

This notebook represents the entire reproduction material required to reproduce and replicate the presented results, tables, and figures as used in the paper "Machine-aided operationalization of continuous software processes".

Please refer to the full technical report [@honel2021technical] for all technical details, code, and data.


```{r}
if (!endsWith(getwd(), "notebooks")) {
  stop("Must be in the notebooks working directory.")
}
```


```{r}
# Available width in inch in the article's template
FIG_WIDTH <- 7.22433 * 0.95 # we'll only occupy 95%
FIG_WIDTH_1COL <- 3.4876

# Common style for all ggplot2 plots
gg2_style <- function(facets = FALSE) {
  sty <- theme_minimal(base_size = 10) + theme(
    text = element_text(family = "serif"),
    axis.title.x.bottom = element_text(margin = margin(t = 8)),
    axis.title.y.left = element_text(margin = margin(r = 8)))
  
  if (facets) {
    sty <- sty + theme(
      text = element_text(family = "LM Roman 10"),
      panel.spacing = unit(1, "mm"),
      panel.border = element_rect(color = "#d8d8d8", fill = NA, size = .1),
      strip.background = element_rect(fill="#f2f2f2", color = "#d8d8d8"),
      strip.text.x = element_text(color="black", margin = margin(t = 2, b = 2)),
      strip.text.y = element_text(color="black", margin = margin(l = 4, r = 2)))
  }
  sty
}
```


# Figures



```{r cont-example-pm, fig.cap="Example of a continuous PM and an observed process."}
plotexpr <- expression({
  set.seed(3)
  
  par(mgp=c(.5,.5,0), mar=c(2,.5,.5,2))

  x <- seq(0, 1, length.out=8)
  l <- loess.smooth(x = x, y = c(1,2.4,3,4,3,3.3,6,5) + rnorm(8, sd=0.3))
  
  
  plot(l$x, l$y - 1, type="l", xlab = NA,
       ylab = NA, xaxt = "n", yaxt = "n", lwd = 4)
  mtext("Time", 1, line = .25, cex=0.8)
  mtext("Amount of activity", side=4, line=.25, cex=0.8)
  grid()
  
  f <- function(x) 2.05+2.2*x - .2*(x+1)^1.7
  l <- loess.smooth(x = x, y = f(x) + rnorm(length(x)))
  lines(l$x, l$y)
  
  legend(0.01, 4.3, legend = c("Process model", "Observed process"), lwd = c(4,1), y.intersp = 1, cex = .7, bg = "white")
})

eval(plotexpr)
```


```{r eval=interactive(), echo=FALSE}
tikzDevice::tikz("../figures/reproduction/cont-example-PM.tex", width = 0.5 * 0.97 * FIG_WIDTH, height = 2)
eval(plotexpr)
dev.off()
```
```{r eval=interactive(), echo=FALSE}
png(filename = "../figures/reproduction/cont-example-PM.png", width = 400, height = 400)
eval(plotexpr)
dev.off()
```




# Research Questions

```{r}
ground_truth <- read.csv(file = "../data/ground-truth.csv", sep = ";")
ground_truth$consensus_score <- ground_truth$consensus / 10
```

## RQ1(b)

Here, we will produce some plots that show the correlation between computed scores.

```{r}
p3_it_scores <- loadResultsOrCompute(file = "../results/p3_it_scores.rds", computeExpr = {
  as.data.frame(
    compute_all_scores_it(
      alignment = p3_it_projects, patternName = "p3_it", vartypes = names(p3_it_signals)))
})
```


```{r per-var-corr-it, fig.cap="Per-variable correlations of computed scores, using issue-tracking data.", fig.height=10, fig.width=10}
temp <- stats::cor(cbind(data.frame(gt_consensus = ground_truth$consensus_score), p3_it_scores[,]))
temp[is.na(temp)] <- 0 # The scores for 'DEV_PEAK' are all 1 and the correlation therefore is NA

corrplot::corrplot(corr = temp, type = "upper",# order = "hclust",
                   tl.col = "black", tl.srt = 90)
```

For issue-tracking data, we have the correlation of scores on a per-variable basis, but not for source code. So next, we will aggregate the scores using the mean across scores. For example, we will produce the average of all `area`-scores.

```{r}
scores <- unique(gsub(pattern = "^(REQ|DEV|DESC)_", replacement = "", x = colnames(p3_it_scores)))
# Remove the mean absolute error, as it's the same as area in our case, since the segment length equaled 1.
scores <- scores[!(scores %in% "mae")]
scores
```

```{r}
p3_it_scores_mean <- data.frame(row.names = rownames(p3_it_scores))

for (name in scores) {
  temp <- cbind(p3_it_scores[, paste0("REQ_", name)], p3_it_scores[, paste0("DEV_", name)], p3_it_scores[, paste0("DESC_", name)])
  temp.mean <- as.vector(apply(X = temp, MARGIN = 1, FUN = mean))
  p3_it_scores_mean <- cbind(p3_it_scores_mean, `colnames<-`(matrix(data = temp.mean, ncol = 1), name))
}


if (interactive()) {
  p3_it_scores_mean
} else {
  knitr::kable(
    x = p3_it_scores_mean,
    booktabs = TRUE,
    caption = "Averaged scores for pattern type 3 (average) for issue-tracking data.",
    label = "p3-it-scores-mean")
}
```

Now with this data, we can show a correlation plot:

```{r corr-it, fig.cap="Correlation of scores using process model type 3 and issue-tracking data."}
plotexpr <- expression({
  par(family = "LM Roman 10")
  temp <- stats::cor(cbind(data.frame(GroundTruth = ground_truth$consensus_score), p3_it_scores_mean[,]))
  temp[is.na(temp)] <- 0 # The scores for 'DEV_PEAK' are all 1 and the correlation therefore is NA
  corrplot::corrplot(corr = temp, type = "upper", tl.col = "black", tl.srt = 90, mar = rep(0,4), tl.cex = 0.85, cl.pos = "b", cl.ratio = .18, cl.length = 5, cl.cex = 1, diag = FALSE, col = colorRampPalette(c("#0037AA", "#eeeeee", "#AA0037"))(100))
})

eval(plotexpr)
```

```{r eval=interactive(), echo=FALSE}
png(filename = "../figures/reproduction/rq1_corr_it.png", width = 400, height = 400)
eval(plotexpr)
invisible(dev.off())
```

```{r}
tikzDevice::tikz("../figures/reproduction/rq1_corr_it.tex", width = 3, height = 2.6 / 2.4 * 3)
eval(plotexpr)
dev.off()
```



Let's load the scores for Source code. Note that these were previously aggregated once using the mean (suffix `_m`) and once using the product (`_p`) across __all__ variables. We will keep the mean for now, and then rename the columns.

```{r}
# P3("avg, no align"actually)
p3_sc_scores <- loadResultsOrCompute(file = "../results/p3_avg_no_scores.rds", computeExpr = {
  as.data.frame(compute_all_scores(alignment = p3_avg_no_align, patternName = "p3_avg"))
})

temp <- NULL
for (name in scores) {
  temp.scores <- `colnames<-`(data.frame(x = p3_sc_scores[, paste0(name, "_m")]), name)
  temp <- if (is.null(temp)) temp.scores else cbind(temp, temp.scores)
}

p3_sc_scores <- temp

if (interactive()) {
  p3_sc_scores
} else {
  knitr::kable(
    x = p3_sc_scores,
    booktabs = TRUE,
    caption = "Scores for pattern type 3 (avg, no align) using source code data.",
    label = "p3-sc-scores")
}
```

```{r corr-sc, fig.cap="Correlation of scores using process model type 3 (avg, no align) and source code data."}
plotexpr <- expression({
  par(family = "LM Roman 10")
  temp <- stats::cor(cbind(data.frame("GroundTruth" = ground_truth$consensus_score), p3_sc_scores[,]))
  temp[is.na(temp)] <- 0 # The scores for 'DEV_PEAK' are all 1 and the correlation therefore is NA
  
  corrplot::corrplot(corr = temp, type = "upper", tl.col = "black", tl.srt = 90, mar = rep(0,4), tl.cex = .85, cl.pos = "b", cl.ratio = .18, cl.length = 5, cl.cex = 1, diag = FALSE, col = colorRampPalette(c("#0037AA", "#eeeeee", "#AA0037"))(100))
})

eval(plotexpr)
```

```{r eval=interactive(), echo=FALSE}
png(filename = "../figures/reproduction/rq1_corr_sc.png", width = 400, height = 400)
eval(plotexpr)
invisible(dev.off())
```

```{r eval=interactive(), echo=FALSE}
tikzDevice::tikz("../figures/reproduction/rq1_corr_sc.tex", width = 3, height = 2.6 / 2.4 * 3)
eval(plotexpr)
dev.off()
```


Since obviously, the metrics have a different correlation between the two types of data, we should compute the pairwise correlations between each score of source code and issue-tracking:

```{r}
temp <- matrix(nrow = 1, ncol = 0)
for (name in scores) {
  temp <- cbind(temp, stats::cor(p3_it_scores_mean[, name], p3_sc_scores[, name]))
}
colnames(temp) <- scores
temp
```

Next, we can examine the correlations between activities. We will do this for issue-tracking data.

```{r}
if (interactive()) {
  p3_it_scores
} else {
  knitr::kable(
    x = p3_it_scores,
    booktabs = TRUE,
    caption = "Per-variable scores for pattern type 3 (average) for issue-tracking data.",
    label = "p3-it-scores")
}
```


## RQ1(c)

We want to produce one large, common correlation-table (not a matrix) using process models 1-3 and 4(1-3).

```{r}
p1_it_signals <- readRDS(file = "../data/p1_it_signals.rds")
p1_it_scores <- readRDS(file = "../results/p1_it_scores.rds")
p2a_it_scores <- readRDS(file = "../results/p2a_it_scores.rds")
p3_it_scores <- readRDS(file = "../results/p3_it_scores.rds")
```


```{r warning=FALSE}
score_types <- c("area", "corr", "jsd", "kl", "arclen", "sd", "var", "mae",
                 "rmse", "RMS", "Kurtosis", "Peak", "ImpulseFactor")
var_types <- names(p1_it_signals)

pAll_it_corr <- matrix(nrow = 3, ncol = length(var_types) * length(score_types))


i <- 0
c_names <- NULL
for (vt in var_types) {
  for (pIdx in 1:3) {
    p_data <- if (pIdx == 1) {
      p1_it_scores } else if (pIdx == 2) {
        p2a_it_scores } else { p3_it_scores }
  
    pAll_it_corr[pIdx, (i*length(score_types)+1):(i*length(score_types)+length(score_types))] <-
      stats::cor(
        x = ground_truth$consensus_score,
        y = p_data[, grepl(pattern = paste0("^", vt), x = colnames(p_data))])
  }
  
  i <- i + 1
  
  c_names <- c(c_names, paste0(vt, "_", score_types))
}

colnames(pAll_it_corr) <- c_names
rownames(pAll_it_corr) <- paste0("PM ", c("I", "II", "III"))
pAll_it_corr[is.na(pAll_it_corr)] <- sqrt(.Machine$double.eps)
```

The average absolute correlation in the table `pAll_it_corr` is __`r round(sum(abs(pAll_it_corr)) / (nrow(pAll_it_corr) * ncol(pAll_it_corr)), 3)`__.

The average absolute correlation per process model is:

```{r}
temp <- apply(pAll_it_corr, 1, function(row) sum(abs(row)) / ncol(pAll_it_corr))

if (interactive()) {
  temp
} else {
  knitr::kable(
    x = temp,
    booktabs = TRUE,
    caption = "The mean absolute correlation per process model (types I--III).",
    label = "mean-abs-corr-pm1-3")
}
```


```{r}
p4p1_it_scores <- readRDS(file = "../results/p4p1_it_scores.rds")
p4p2a_it_scores <- readRDS(file = "../results/p4p2a_it_scores.rds")
p4p3_it_scores <- readRDS(file = "../results/p4p3_it_scores.rds")
```


```{r}
p4All_it_corr <- matrix(nrow = 3, ncol = length(var_types) * length(score_types))

i <- 0
c_names <- NULL
for (vt in var_types) {
  for (pIdx in 1:3) {
    p_data <- if (pIdx == 1) {
      p4p1_it_scores } else if (pIdx == 2) {
        p4p2a_it_scores } else { p4p3_it_scores }
  
    p4All_it_corr[pIdx, (i*length(score_types)+1):(i*length(score_types)+length(score_types))] <-
      stats::cor(
        x = ground_truth$consensus_score,
        y = p_data[, grepl(pattern = paste0("^", vt), x = colnames(p_data))])
  }
  
  i <- i + 1
  
  c_names <- c(c_names, paste0(vt, "_", score_types))
}

colnames(p4All_it_corr) <- c_names
rownames(p4All_it_corr) <- paste0("PM IV(", 1:3, ")")
p4All_it_corr[is.na(p4All_it_corr)] <- sqrt(.Machine$double.eps)
```

The average absolute correlation in the table `p4All_it_corr` is __`r round(sum(abs(p4All_it_corr)) / (nrow(p4All_it_corr) * ncol(p4All_it_corr)), 3)``__.

The average absolute correlation per process model is:

```{r}
temp <- apply(p4All_it_corr, 1, function(row) sum(abs(row)) / ncol(p4All_it_corr))

if (interactive()) {
  temp
} else {
  knitr::kable(
    x = temp,
    booktabs = TRUE,
    caption = "The mean absolute correlation per derivative process model (types IV(I--III)).",
    label = "mean-abs-corr-pm1-3")
}
```

```{r corr-all-it, fig.cap="Per-variable and per-process model correlations of scoresusing issue-tracking data.", message=FALSE}
plotexpr <- expression({
  temp <- t(rbind(
    `rownames<-`(x = pAll_it_corr, value = c("I", "II", "III")),
    `rownames<-`(x = p4All_it_corr, value = c("*I", "*II", "*III"))
  ))
  temp <- temp[, rev(colnames(temp))]
  
  rownames(temp) <- sapply(X = rownames(temp), function(n) {
    sp <- strsplit(x = n, split = "_")[[1]]
    paste0("$\\mathsf{", sp[1], "}_{\\mathrm{", sp[2], "}}$")
  })
  
  
  ggcorrplot::ggcorrplot(corr = temp, ggtheme = gg2_style() + theme(
      axis.text.x.top = element_text(angle = 90, vjust = 1/3, hjust = 0, size = 8),
      axis.text.y.left = element_text(size = 8),
      legend.position = "bottom",
      legend.text = element_text(vjust = 0),
      legend.title = element_text(margin = margin(r = 10, b = 3)),
      legend.key.width = unit(20, 'pt'),
      legend.key.height = unit(8, 'pt'),
      axis.title.y.left = element_blank())) +
  scale_x_discrete(position = "top") +
  scale_fill_gradient2(
    low = "#0037AA", mid = "#eeeeee", high = "#AA0037", guide = "colorbar",
    limits = c(-1,1), breaks = seq(-1, 1, by = .5)) +
  labs(fill = "Pearson sample correlation")
})

eval(plotexpr)
```
```{r eval=interactive(), echo=FALSE}
tikzDevice::tikz("../figures/reproduction/rq1_corr_all.tex", width = FIG_WIDTH, height = 2.25)
eval(plotexpr)
dev.off()
```



# References {-}

<div id="refs"></div>

