---
title: "Technical Reports Compilation: Detecting the Fire Drill Anti-pattern Using Source Code and Issue-Tracking Data"
author: "Sebastian Hönel"
date: "`r format(Sys.time(), '%B %d, %Y')`"
bibliography: ../inst/REFERENCES.bib
urlcolor: blue
output:
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 6
    lof: yes
    lot: yes
    df_print: kable
    keep_tex: yes
  bookdown::pdf_book:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 6
    df_print: kable
    keep_tex: yes
  md_document:
    toc: yes
    toc_depth: 6
    df_print: kable
    variant: gfm
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 6
    toc_float: yes
    df_print: kable
  word_document: default
  pdf_document:
    toc: yes
    toc_depth: '6'
header-includes:
- \usepackage{bm}
- \usepackage{mathtools}
- \usepackage{xurl}
abstract: "\\noindent Detecting the presence of project management anti-patterns (AP) currently requires experts on the matter and is an expensive endeavor. Worse, experts may introduce their individual subjectivity or bias. Using the Fire Drill AP, we first introduce a novel way to translate descriptions into detectable AP that are comprised of arbitrary metrics and events such as logged time or maintenance activities, which are mined from the underlying source code or issue-tracking data, thus making the description objective as it becomes data-based. Secondly, we demonstrate a novel method to quantify and score the deviations of real-world projects to data-based AP descriptions. Using fifteen real-world projects that exhibit a Fire Drill to some degree, we show how to further enhance the translated AP. The ground truth in these projects was extracted from two individual experts and consensus was found between them. We introduce a novel method called automatic calibration, that optimizes a pattern such that only necessary and important scores remain that suffice to confidently detect the degree to which the AP is present. Without automatic calibration, the proposed patterns show only weak potential for detecting the presence. Enriching the AP with data from real-world projects significantly improves the potential. We also introduce a no-pattern approach that exploits the ground truth for establishing a new, quantitative understanding of the phenomenon, as well as for finding gray-/black-box predictive models. We conclude that the presence detection and severity assessment of the Fire Drill anti-pattern, as well as some of its related and similar patterns, is certainly possible using some of the presented approaches."
---

\newcommand*\mean[1]{\overline{#1}}
\newcommand{\abs}[1]{\left\lvert\,#1\,\right\rvert}
\newcommand{\norm}[1]{\left\lVert\,#1\,\right\rVert}
\newcommand{\infdiv}[2]{#1\;\|\;#2}
\newcommand\argmax[1]{\underset{#1}{arg\,max}}
\newcommand\argmin[1]{\underset{#1}{arg\,min}}

```{r setoptions, echo=FALSE, warning=FALSE, message=FALSE}
Sys.setenv(LANG = "en_US.UTF-8")
Sys.setenv(LC_ALL = "en_US.UTF-8")
Sys.setenv(LC_CTYPE = "en_US.UTF-8")
library(knitr)
opts_chunk$set(tidy = TRUE, tidy.opts = list(indent=2))
```

```{r echo=FALSE, warning=FALSE}
source(file = "../helpers_rank-model.R")
source(file = "../models/rank-model.R")
source(file = "../helpers.R")
source(file = "./common-funcs.R")
source(file = "../models/modelsR6.R")
source(file = "../models/SRBTW-R6.R")

library(ggplot2)
library(ggpubr)
```

# Overview

This document is a compilation of five separate technical reports.
The canonical reference to this report is [@honel2021technical].
In all detail, the development of methods for detecting the presence of so-called _\"anti-patterns\"_ in software development projects is presented.

The first technical report is concerned with this concrete problem, and it facilitates two major building blocks: The first is the application of a new method for time warping, called self-regularizing boundary time/amplitude warping (srBTAW). The second building block is a detailed walk-through of creating a classifier for commits, based on source code density. Both these blocks have dedicated technical reports.

The second technical report is concerned with the same problem, but it facilitates issue-tracking data, as well as additional methods for detecting, such as inhomogeneous confidence intervals and vector fields.

The third technical report is an attempt to replace the approach of unsupervised scoring with a robust regression model that has good generalizability using stable learning through adaptive training.
Also, it add a slight adaptation of the first source code pattern.

Notebooks four and five are concerned with an optimization framework for curve warping and -transformation, as well as classifying commits, respectively.

All of the data, source code, and raw materials can be found online. These reports and resources are made available for reproduction purposes. The interested reader is welcome and enabled to re-run all of the computations and to extend upon our ideas. The repository is to be found at <https://github.com/MrShoenel/anti-pattern-models>. The data is made available online [@honel_picha_2021].
